---
#HTML DOC
title: "RSA_Graphs"
author: "Estibaliz Herrera"
date: "16-07-2026"
output:
  html_document:
    df_print: paged
---
# WORD DOC
# title: "RSA_Graphs"
# author: "Estibaliz Herrera"
# date: "16-07-2026"
# output:
#   html_document:
#     df_print: paged
#   word_document: default
---

```{r setup, include=FALSE}
#DEBUGGING MODE:
# knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)

# KNITTING MODE: (WHEN DONE DEBUGGING)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
#install
#install.packages("plotly")
#install.packages("paletteer")
#install.packages("colorspace")
#install.packages("readr")
#install.packages("afex")
#install.packages("emmeans")
#install.packages("tibble")
#install.packages("scales")
#install.packages("ggsci")
#install.packages("MetBrewer") 
#install.packages("effectsize")
#install.packages("ggtext")
#install.packages("Hmisc")
#install.packages("NatParksPalettes")
#install.packages("conflicted")  
```

```{r}
# library(plyr) # plyr conflicts with lots of other packages; just call it when you need it
library(tidyverse)
library(ggplot2) 
library(glue)
library(grid)
library(png)
library(car)
library(tibble)
library(afex)
library(emmeans)
library(effectsize)
library(viridis)
library(ggsci)  # https://nanx.me/ggsci/
library(plotly) # creates interactive graphics (needs ggplot)
library(paletteer) #palette finder: https://r-graph-gallery.com/color-palette-finder.html
library(colorspace)
library(scales)
library(ggsci)
library(MetBrewer)
library(NatParksPalettes)
library(colorspace)
library(conflicted)

# Potential Library Conflicts
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::mutate)
conflicts_prefer(dplyr::summarise)

```
# This chunk will work globally for all ANOVAs
```{r}

# ---- GLOBAL OPTIONS ----
options(contrasts = c("contr.sum", "contr.poly"))

```

```{r}

# Setting the Path

library(readr)

datapath <- "C:/Users/ellave3/OneDrive - Georgia Institute of Technology/Documents/GitHub/OLNOLTimecourses/RSA_graphs/sourceData"

filenames <- list.files(
  path = datapath,
  pattern = "^matlab_rsa_spreadsheet_RAW.*\\.csv$",
  full.names = TRUE
)

stopifnot(length(filenames) > 0)

newest_file <- filenames[which.max(file.info(filenames)$mtime)]

data.raw <- read_csv(newest_file, show_col_types = FALSE) #grabs the most up to date spreadsheet

```
# DATA CLEANING AND RELABELING
```{r}

# 1) pivoting > long format

long_raw <- data.raw %>%
  pivot_longer(
    cols = -ID,
    names_to = "raw_name",
    values_to = "RSA"
  )

# 2) Parse variables

long_parsed <- long_raw %>%
  mutate(
    ROI = case_when(
      str_detect(raw_name, "hippo") ~ "Hippocampus",
      str_detect(raw_name, "RSComplex") ~ "Complex",
      str_detect(raw_name, "ba2930") ~ "RSC"
    ),
    
    learningStage = case_when(
      str_detect(raw_name, "_early_") ~ "Early",
      str_detect(raw_name, "_late_") ~ "Late"
    ),
  
    Hall = case_when(
      str_detect(raw_name, "_0") ~ "Cue period",
      str_detect(raw_name, "_1") ~ "1st hallway",
      str_detect(raw_name, "_2") ~ "2nd hallway",
      str_detect(raw_name, "_3") ~ "Goal"
   ),
  
Overlappicity = case_when(
  str_detect(raw_name, "^'?nol_") ~ "NOL",
  str_detect(raw_name, "^'?ol_")  ~ "OL"
  ),
  
  
  )



```
# FIX FACTOR TYPES & LEVELS ONCE, GLOBALLY 
```{r}
# Data reorganization: coverting learningStage into a factor (needs to be read by aov_ez)
# ---- CANONICAL FACTOR NORMALIZATION ----
long_parsed <- long_parsed %>%
  mutate(
    learningStage = factor(learningStage, levels = c("Early", "Late")),
    Hall = factor(
      Hall,
      levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")
    ),
    ROI = factor(
      ROI,
      levels = c("Hippocampus", "RSC", "Complex")
    ),
    Overlappicity = factor(Overlappicity, levels = c("OL", "NOL"))
  )

# Sanity check (ensuring downstream chunks aren't fighting corrupted data)
stopifnot(
  is.factor(long_parsed$Hall),
  is.factor(long_parsed$ROI),
  is.factor(long_parsed$learningStage)
)

```

preview of color palettes
```{r}

# library(paletteer)
# library(scales) # for show_col()

# Purples & Pinks
# paletteer::paletteer_d("RColorBrewer::PuRd") |> scales::show_col() # Orchid, fucsias...
# paletteer::paletteer_d("RColorBrewer::Purples") |> scales::show_col() # cool toned (grayish) purples
# paletteer::paletteer_d("RColorBrewer::RdPu") |> scales::show_col() # pinks → purple

# Purples / Warm violets
# paletteer::paletteer_d("RColorBrewer::Purples") |> scales::show_col()  # light → dark purple
# paletteer::paletteer_d("RColorBrewer::RdPu")   |> scales::show_col()   # red → purple
# paletteer::paletteer_d("RColorBrewer::PuRd")   |> scales::show_col()   # purple → red
# paletteer::paletteer_d("RColorBrewer::BuPu")  |> scales::show_col() # blue-purples

# Reds & Oranges
# paletteer::paletteer_d("RColorBrewer::YlOrBr") |> scales::show_col() # oranges → yellows
# paletteer::paletteer_d("RColorBrewer::YlOrRd") |> scales::show_col()   # yellow → dark red
# paletteer::paletteer_d("RColorBrewer::Oranges") |> scales::show_col()  # light → dark orange

# Combined warm oranges 
# paletteer::paletteer_d("RColorBrewer::YlOrBr") |> scales::show_col()  # yellow → orange → dark brownish red

# Blues
# paletteer::paletteer_d("RColorBrewer::Blues") |> scales::show_col() # blues
# paletteer::paletteer_d("RColorBrewer::PuBuGn") |> scales::show_col()  # purple → teal → blue

# Greens
# paletteer::paletteer_d("RColorBrewer::Greens") |> scales::show_col() # greens
# paletteer::paletteer_d("RColorBrewer::GnBu")   |> scales::show_col() # blue-greens
# paletteer::paletteer_d("RColorBrewer::YlGn")  |> scales::show_col() # yellow-greens

#Blues II (nice for publications)
# paletteer::paletteer_d("MetBrewer::Hiroshige") |> scales::show_col()

## Correlation Matrix Palettes
#scales::show_col(NatParksPalettes::natparks.pals("Acadia", 25, type = "continuous"), labels = TRUE)
#NatParksPalettes::display_all()
#colours = viridisLite::viridis(256),  # "viridis", "magma", "plasma", "cividis"
  # colours = rev(viridisLite::viridis(256)), # I like this one, looks readable
  #colours = rev(NatParksPalettes::natparks.pals("Acadia", 256, type = "continuous")),
  #colours = (NatParksPalettes::natparks.pals("SouthDowns", 256, type = "continuous")), # I like this one, looks readable
  ## colours = rev(NatParksPalettes::natparks.pals("Olympic", 256, type = "continuous")), # I like this one, looks readable

  #POTENTIAL ALTERNATIVE TO THE (REV) VIRIDIS LIBRARY I am using
  # scale_fill_gradient2( # Sets the color scale for fill
  #   low = "white", # for a 0-centered diverging scale with a few negative values, set a light color for negatives
  #   high = "blue",
  #   na.value = "white",
  #   midpoint = 0, # centers the color scale at 0 (useful for correlations)
  #   limits = c(-0.2, 1),
  #   name = expression(italic(r)) # legend title
  # )

```
PLOTS COLOR CODE

```{r}
# 1) ROI colors

# 2a) PLOTS
HC_OL_Early_orange <- paletteer::paletteer_d("RColorBrewer::YlOrBr", n = 9)[5]
HC_OL_Late_orange <- paletteer::paletteer_d("RColorBrewer::YlOrBr", n = 9)[6]
RSC_OL_Early_purple <- paletteer::paletteer_d("RColorBrewer::RdPu", n = 9)[8]
RSC_OL_Late_purple <- paletteer::paletteer_d("RColorBrewer::RdPu", n = 9)[7]
Complex_OL_Early_blue <- paletteer::paletteer_d("MetBrewer::Hiroshige", n = 10)[7] #changed but not sure it'll pick up #7AA6DCFF
Complex_OL_Late_blue <- paletteer::paletteer_d("MetBrewer::Hiroshige", n = 10)[9] #changed but not sure 

HC_NOL_Early_orange <- paletteer::paletteer_d("RColorBrewer::Oranges", n = 9)[6] #needs to change 
HC_NOL_Late_orange <- paletteer::paletteer_d("RColorBrewer::YlOrBr", n = 9)[7] 
RSC_NOL_Early_purple <- paletteer::paletteer_d("RColorBrewer::Purples", n = 9)[7]
RSC_NOL_Late_purple <- paletteer::paletteer_d("RColorBrewer::BuPu", n = 9)[8]
Complex_NOL_Early_blue <- paletteer::paletteer_d("RColorBrewer::Blues", n = 9)[5] 
Complex_NOL_Late_blue <- paletteer::paletteer_d("RColorBrewer::Blues", n = 9)[8]

```
****GRAPHS****
***************************
1) “how much does the representation of the environments change within a given ROI with learning [EARLY vs LATE]?” 
***********************************
ROI: EARLY vs LATE
**************************
HIPPOCAMPUS OL
**************************
-> Repeated measures ANOVA Learning stage (Early,Late) x Hall (Cue period, 1st hall, 2nd hall, goal)
(1) OL HC
```{r}

# 1) Filtering + Subject means
plot_data_OL_HC <- long_parsed %>%
  filter(
    ROI %in% c("Hippocampus"),     # ROI filter
    str_detect(raw_name, "ol_new_early_w_ol_new_early_|ol_new_late_w_ol_new_late_") 
  ) %>%
  group_by(ID, learningStage, Hall) %>%
  summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop") #%>%
  # 
  # # Ensure Hall is a factor in the correct order
  # mutate(Hall = factor(Hall,
  #                      levels = c("Cue period",
  #                                 "1st hallway",
  #                                 "2nd hallway",
  #                                 "Goal")))

  
  
# ✅ STOP PIPE HERE
stopifnot(nrow(plot_data_OL_HC) > 0)

print(plot_data_OL_HC)

```
ANOVA
```{r}

# summary
# I'll have to relabel 'anova_model', 'anova_table', etc if I don't want these data to be overwritten
# options(contrasts = c("contr.sum", "contr.poly")) #same analysis as in JASP
anova_model_OL_HC <- aov_ez(
  id = "ID",
  dv = "RSA",
  data = plot_data_OL_HC,
  within = c("learningStage", "Hall"),
  type = 3, # Type III SS (JASP style)
  anova_table = list(
    correction = "GG", # <-- Greenhouse–Geisser correction for when factors ≥3 levels (e.g., Halls)
    es = "pes"   # partial eta squared
  )
)


# Extract ANOVA table as data frame
anova_table_OL_HC <- as.data.frame(anova_model_OL_HC$anova_table) |>
  tibble::rownames_to_column(var = "Effect")

 
# Format p-values like JASP
format_p <- function(p) {
  if (p < .001) "< .001"
  else paste0("= ", format(round(p, 3), nsmall = 3))
}

 # format_p <- function(p) {
 #   if (p < .001) {
 #     "< .001"
 #  } else {
 #     paste0("= ", format(round(p, 3), nsmall = 3))
 #   }
 # }

# Clean ANOVA table
niceAnovaTable_OL_HC <- data.frame(
  Effect = anova_table_OL_HC$Effect,
  df1    = anova_table_OL_HC[[1]],
  df2    = anova_table_OL_HC[[2]],
  F      = round(anova_table_OL_HC$F, 3),
  p      = sapply(anova_table_OL_HC$`Pr(>F)`, format_p),
  `η²p`  = round(anova_table_OL_HC$pes, 3)
)

print(niceAnovaTable_OL_HC)


```
POST-HOCs 
```{r}

# Estimated marginal means for ALL 6 estimates (avg over 'learningStage')
emm_Hall_OL_HC <- emmeans(anova_model_OL_HC, ~ Hall)


# Holm correction
posthoc_OL_HC <- pairs(emm_Hall_OL_HC, adjust = "holm")

holm_table_OL_HC <- summary(posthoc_OL_HC) |>
  as.data.frame() |>
  mutate(
    estimate = round(estimate, 3),
    SE       = round(SE, 3),
    t        = round(t.ratio, 2),

    ##--------->  Incorrect RM Cohen’s d
    cohen_d  = round(abs(t.ratio) / sqrt(df + 1), 3),
    
     pHolms = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ paste0("= ", sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3))))
    ),

    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE ~ ""
    )
   
  )

print(holm_table_OL_HC)


```
DESCRIPTIVES

```{r}
###) Descriptives
descriptives_OL_HC <- plot_data_OL_HC %>%
  group_by(learningStage, Hall) %>%
  summarise(
    N = n(),
    Mean = mean(RSA, na.rm = TRUE),
    SD = sd(RSA, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(across(c(Mean, SD, SE), ~ round(.x, 3)))

print(descriptives_OL_HC)
```
PLOTTING OL HIPPOCAMPUS (LearningStage x Hall)

```{r}

# 2) Plot with ribbon for mean ± SE
p_OL_HC <- ggplot(plot_data_OL_HC,
       aes(x = Hall,
           y = RSA,
           color = learningStage,
           fill = learningStage,
           group = learningStage)) +
  
  # Ribbon for standard error
  stat_summary(fun.data = mean_se,
               geom = "ribbon",
               alpha = 0.15,
               color = NA) +
  
   # Mean line with Early dashed, Late solid
  stat_summary(fun = mean,
               geom = "line",
               aes(linetype = learningStage),
               linewidth = 1) +
  
  # Mean solid line
  # stat_summary(fun = mean,
  #              geom = "line",
  #              linewidth = 1) +
  
  # Mean points
  stat_summary(fun = mean,
               geom = "point",
               size = 3) +
  
  # Custom colors
  
  scale_color_manual(values = c(
    "Early" = HC_OL_Early_orange,
    "Late" = HC_OL_Late_orange
  )) +
  
  scale_fill_manual(values = c(
   "Early" = HC_OL_Early_orange,
    "Late" = HC_OL_Late_orange
  )) +
  
  # Line types
  scale_linetype_manual(values = c(
    "Early" = "dashed",
    "Late" = "solid"
  )) +
  
   # Y-axis formatting ONLY
  scale_y_continuous(
    breaks = seq(0, 0.12, by = 0.02),
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
  ) +
  
  # Zoom display WITHOUT removing data
  coord_cartesian(
     ylim = c(0, 0.12)
  ) + 
  
  # Labels
  labs(x = "Hall",
       y = "Representational similarity (r)",
       title = "(1) OL Learning Stage HC",
       # subtitle = "(1) xxx",
       color = "Learning stage",
       fill = "Learning stage",
       linetype = "Learning stage") +
  
  # Clean theme
  theme_classic(base_size = 14)

print(p_OL_HC)

# Save graph
ggsave(
  "OL_Early-Late_HC.tiff",
  plot = p_OL_HC,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)


```

**************************
HIPPOCAMPUS NOL
**************************

-> Repeated measures ANOVA Learning stage (Early,Late) x Hall (Cue period, 1st hall, 2nd hall, goal)
(1) NOL HC 
```{r}

# 1) Filtering + Subject means
plot_data_NOL_HC <- long_parsed %>%
  filter(
    ROI %in% c("Hippocampus"),     # ROI filter
    str_detect(raw_name, "nol_new_early_w_nol_new_early_|nol_new_late_w_nol_new_late_") 
  ) %>%
  group_by(ID, learningStage, Hall) %>%
  summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop") #%>%
  # 
  # # Ensure Hall is a factor in the correct order
  # mutate(Hall = factor(Hall,
  #                      levels = c("Cue period",
  #                                 "1st hallway",
  #                                 "2nd hallway",
  #                                 "Goal")))

  
  
# ✅ STOP PIPE HERE
stopifnot(nrow(plot_data_NOL_HC) > 0)

print(plot_data_NOL_HC)

```
ANOVA
```{r}

# summary
# I'll have to relabel 'anova_model', 'anova_table', etc if I don't want these data to be overwritten
# options(contrasts = c("contr.sum", "contr.poly")) #same analysis as in JASP
anova_model_NOL_HC <- aov_ez(
  id = "ID",
  dv = "RSA",
  data = plot_data_NOL_HC,
  within = c("learningStage", "Hall"),
  type = 3, # Type III SS (JASP style)
  anova_table = list(
    correction = "GG", # <-- Greenhouse–Geisser correction for when factors ≥3 levels (e.g., Halls)
    es = "pes"   # partial eta squared
  )
)


# Extract ANOVA table as data frame
anova_table_NOL_HC <- as.data.frame(anova_model_NOL_HC$anova_table) |>
  tibble::rownames_to_column(var = "Effect")

 
# Format p-values like JASP
format_p <- function(p) {
  if (p < .001) "< .001"
  else paste0("= ", format(round(p, 3), nsmall = 3))
}

 # format_p <- function(p) {
 #   if (p < .001) {
 #     "< .001"
 #  } else {
 #     paste0("= ", format(round(p, 3), nsmall = 3))
 #   }
 # }

# Clean ANOVA table
niceAnovaTable_NOL_HC <- data.frame(
  Effect = anova_table_NOL_HC$Effect,
  df1    = anova_table_NOL_HC[[1]],
  df2    = anova_table_NOL_HC[[2]],
  F      = round(anova_table_NOL_HC$F, 3),
  p      = sapply(anova_table_NOL_HC$`Pr(>F)`, format_p),
  `η²p`  = round(anova_table_NOL_HC$pes, 3)
)

print(niceAnovaTable_NOL_HC)


```
POST-HOCs 
```{r}

# no main effects nor interactions: no need to report this
# Estimated marginal means for ALL 6 estimates (avg over 'learningStage')
emm_Hall_NOL_HC <- emmeans(anova_model_NOL_HC, ~ Hall)


# Holm correction
posthoc_NOL_HC <- pairs(emm_Hall_NOL_HC, adjust = "holm")

holm_table_NOL_HC <- summary(posthoc_NOL_HC) |>
  as.data.frame() |>
  mutate(
    estimate = round(estimate, 3),
    SE       = round(SE, 3),
    t        = round(t.ratio, 2),

    ##--------->  Incorrect RM Cohen’s d
    cohen_d  = round(abs(t.ratio) / sqrt(df + 1), 3),
    
     pHolms = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ paste0("= ", sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3))))
    ),

    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE ~ ""
    )
   
  )

print(holm_table_NOL_HC)


```
DESCRIPTIVES

```{r}
###) Descriptives
descriptives_NOL_HC <- plot_data_NOL_HC %>%
  group_by(learningStage, Hall) %>%
  summarise(
    N = n(),
    Mean = mean(RSA, na.rm = TRUE),
    SD = sd(RSA, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(across(c(Mean, SD, SE), ~ round(.x, 3)))

print(descriptives_NOL_HC)
```
PLOTTING NOL HIPPOCAMPUS (LearningStage x Hall)

```{r}

# 2) Plot with ribbon for mean ± SE
p_NOL_HC <- ggplot(plot_data_NOL_HC,
       aes(x = Hall,
           y = RSA,
           color = learningStage,
           fill = learningStage,
           group = learningStage)) +
  
  # Ribbon for standard error
  stat_summary(fun.data = mean_se,
               geom = "ribbon",
               alpha = 0.15,
               color = NA) +
  
   # Mean line with Early dashed, Late solid
  stat_summary(fun = mean,
               geom = "line",
               aes(linetype = learningStage),
               linewidth = 1) +
  
  # Mean solid line
  # stat_summary(fun = mean,
  #              geom = "line",
  #              linewidth = 1) +
  
  # Mean points
  stat_summary(fun = mean,
               geom = "point",
               size = 3) +
  
  # Custom colors
  
  scale_color_manual(values = c(
    "Early" = HC_NOL_Early_orange,
    "Late" = HC_NOL_Late_orange
  )) +
  
  scale_fill_manual(values = c(
   "Early" = HC_NOL_Early_orange,
    "Late" = HC_NOL_Late_orange
  )) +
  
  # Line types
  scale_linetype_manual(values = c(
    "Early" = "dashed",
    "Late" = "solid"
  )) +
  
   # Y-axis formatting ONLY
  scale_y_continuous(
    breaks = seq(0, 0.12, by = 0.02),
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
  ) +
  
  # Zoom display WITHOUT removing data
  coord_cartesian(
     ylim = c(0, 0.12)
  ) + 
  
  # Labels
  labs(x = "Hall",
       y = "Representational similarity (r)",
       title = "(1) NOL Learning Stage HC",
       # subtitle = "(2a) xxx",
       color = "Learning stage",
       fill = "Learning stage",
       linetype = "Learning stage") +
  
  # Clean theme
  theme_classic(base_size = 14)

print(p_NOL_HC)

# Save graph
ggsave(
  "NOL_Early-Late_HC.tiff",
  plot = p_NOL_HC,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)


```

**************************
RSC OL
**************************

-> Repeated meausures ANOVA Learning stage (Early,Late) x Hall (Cue period, 1st hall, 2nd hall, goal)
(1) OL RSC 
```{r}
# 1) Filtering + Subject means
plot_data_OL_RSC <- long_parsed %>%
  filter(
    ROI %in% c("RSC"), 
    str_detect(raw_name, "ol_new_early_w_ol_new_early_|ol_new_late_w_ol_new_late_") 
  ) %>%
  group_by(ID, learningStage, Hall) %>%
  summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop")

# ✅ STOP PIPE HERE
stopifnot(nrow(plot_data_OL_RSC) > 0)

print(plot_data_OL_RSC)

```
ANOVA
```{r}
anova_model_OL_RSC <- aov_ez(
  id = "ID",
  dv = "RSA",
  data = plot_data_OL_RSC,
  within = c("learningStage", "Hall"),
  type = 3,
  anova_table = list(
    correction = "GG",
    es = "pes"
  )
)

# Extract ANOVA table
anova_table_OL_RSC <- as.data.frame(anova_model_OL_RSC$anova_table) |>
  tibble::rownames_to_column(var = "Effect")

# Format p-values
format_p <- function(p) {
  if (p < .001) "< .001"
  else paste0("= ", format(round(p, 3), nsmall = 3))
}

# Clean table
niceAnovaTable_OL_RSC <- data.frame(
  Effect = anova_table_OL_RSC$Effect,
  df1    = anova_table_OL_RSC[[1]],
  df2    = anova_table_OL_RSC[[2]],
  F      = round(anova_table_OL_RSC$F, 3),
  p      = sapply(anova_table_OL_RSC$`Pr(>F)`, format_p),
  `η²p`  = round(anova_table_OL_RSC$pes, 3)
)

print(niceAnovaTable_OL_RSC)

```
POST-HOCs 
```{r}

# EMMs averaged over learningStage (6 comparisons)
emm_Hall_OL_RSC <- emmeans(anova_model_OL_RSC, ~ Hall)

# Holm correction
posthoc_OL_RSC <- pairs(emm_Hall_OL_RSC, adjust = "holm")

holm_table_OL_RSC <- summary(posthoc_OL_RSC) |>
  as.data.frame() |>
  mutate(
    estimate = round(estimate, 3),
    SE       = round(SE, 3),
    t        = round(t.ratio, 2),

    # Cohen’s d
    cohen_d  = round(abs(t.ratio) / sqrt(df + 1), 3),

    pHolms = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ paste0("= ", sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3))))
    ),

    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE ~ ""
    )
  )

print(holm_table_OL_RSC)

```
DESCRIPTIVES

```{r}
descriptives_OL_RSC <- plot_data_OL_RSC %>%
  group_by(learningStage, Hall) %>%
  summarise(
    N = n(),
    Mean = mean(RSA, na.rm = TRUE),
    SD = sd(RSA, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(across(c(Mean, SD, SE), ~ round(.x, 3)))

print(descriptives_OL_RSC)

```
PLOTTING OL RSC (LearningStage x Hall)

```{r}

# 2) Plot with ribbon for mean ± SE
p_OL_RSC <- ggplot(plot_data_OL_RSC,
       aes(x = Hall,
           y = RSA,
           color = learningStage,
           fill = learningStage,
           group = learningStage)) +
  
  # Ribbon for standard error
  stat_summary(fun.data = mean_se,
               geom = "ribbon",
               alpha = 0.15,
               color = NA) +
  
# Mean line with Early dashed, Late solid
  stat_summary(fun = mean,
               geom = "line",
               aes(linetype = learningStage),
               linewidth = 1) +
  
  # Mean points
  stat_summary(fun = mean,
               geom = "point",
               size = 3) +
  
  # Custom colors
  
  scale_color_manual(values = c(
    "Early" = RSC_OL_Early_purple,
    "Late" = RSC_OL_Late_purple
  )) +
  
  scale_fill_manual(values = c(
   "Early" = RSC_OL_Early_purple,
    "Late" = RSC_OL_Late_purple
  )) +
  
  # Line types
  scale_linetype_manual(values = c(
    "Early" = "dashed",
    "Late" = "solid"
  )) +
  
   # Y-axis formatting ONLY
  scale_y_continuous(
    breaks = seq(0, 0.12, by = 0.02),
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
  ) +
  
  # Zoom display WITHOUT removing data
  coord_cartesian(
     ylim = c(0, 0.12)
  ) + 
  
   # Labels
  labs(x = "Hall",
       y = "Representational similarity (r)",
       title = "(1) OL Learning Stage RSC",
       color = "Learning stage",
       fill = "Learning stage",
       linetype = "Learning stage") +
  
  
  # Clean theme
  theme_classic(base_size = 14)

print(p_OL_RSC)

# Save graph
ggsave(
  "OL_Early-Late_RSC.tiff",
  plot = p_OL_RSC,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)


```

**************************
RSC NOL
**************************

-> Repeated meausures ANOVA Learning stage (Early,Late) x Hall (Cue period, 1st hall, 2nd hall, goal)
(1) NOL RSC 
```{r}

# 1) Filtering + Subject means
plot_data_NOL_RSC <- long_parsed %>%
  filter(
    ROI %in% c("RSC"),     
    str_detect(raw_name, "nol_new_early_w_nol_new_early_|nol_new_late_w_nol_new_late_") 
  ) %>%
  group_by(ID, learningStage, Hall) %>%
  summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop")

# ✅ STOP PIPE HERE
stopifnot(nrow(plot_data_NOL_RSC) > 0)

print(plot_data_NOL_RSC)

```
ANOVA
```{r}
anova_model_NOL_RSC <- aov_ez(
  id = "ID",
  dv = "RSA",
  data = plot_data_NOL_RSC,
  within = c("learningStage", "Hall"),
  type = 3,
  anova_table = list(
    correction = "GG",
    es = "pes"
  )
)

# Extract ANOVA table
anova_table_NOL_RSC <- as.data.frame(anova_model_NOL_RSC$anova_table) |>
  tibble::rownames_to_column(var = "Effect")

# Format p-values
format_p <- function(p) {
  if (p < .001) "< .001"
  else paste0("= ", format(round(p, 3), nsmall = 3))
}

# Clean table
niceAnovaTable_NOL_RSC <- data.frame(
  Effect = anova_table_NOL_RSC$Effect,
  df1    = anova_table_NOL_RSC[[1]],
  df2    = anova_table_NOL_RSC[[2]],
  F      = round(anova_table_NOL_RSC$F, 3),
  p      = sapply(anova_table_NOL_RSC$`Pr(>F)`, format_p),
  `η²p`  = round(anova_table_NOL_RSC$pes, 3)
)

print(niceAnovaTable_NOL_RSC)

```
POST-HOCs 
```{r}

# EMMs averaged over learningStage (6 comparisons)
emm_Hall_NOL_RSC <- emmeans(anova_model_NOL_RSC, ~ Hall)

# Holm correction
posthoc_NOL_RSC <- pairs(emm_Hall_NOL_RSC, adjust = "holm")

holm_table_NOL_RSC <- summary(posthoc_NOL_RSC) |>
  as.data.frame() |>
  mutate(
    estimate = round(estimate, 3),
    SE       = round(SE, 3),
    t        = round(t.ratio, 2),

    # ✅ RM Cohen’s d
    cohen_d  = round(abs(t.ratio) / sqrt(df + 1), 3),

    pHolms = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ paste0("= ", sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3))))
    ),

    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE ~ ""
    )
  )

print(holm_table_NOL_RSC)

```
DESCRIPTIVES

```{r}
descriptives_NOL_RSC <- plot_data_NOL_RSC %>%
  group_by(learningStage, Hall) %>%
  summarise(
    N = n(),
    Mean = mean(RSA, na.rm = TRUE),
    SD = sd(RSA, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(across(c(Mean, SD, SE), ~ round(.x, 3)))

print(descriptives_NOL_RSC)
```
PLOTTING NOL RSC (LearningStage x Hall)

```{r}

# 2) Plot with ribbon for mean ± SE
p_NOL_RSC <- ggplot(plot_data_NOL_RSC,
       aes(x = Hall,
           y = RSA,
           color = learningStage,
           fill = learningStage,
           group = learningStage)) +
  
  # Ribbon for standard error
  stat_summary(fun.data = mean_se,
               geom = "ribbon",
               alpha = 0.15,
               color = NA) +
  
# Mean line with Early dashed, Late solid
  stat_summary(fun = mean,
               geom = "line",
               aes(linetype = learningStage),
               linewidth = 1) +
  
  # Mean points
  stat_summary(fun = mean,
               geom = "point",
               size = 3) +
  
  # Custom colors
  
  scale_color_manual(values = c(
    "Early" = RSC_NOL_Early_purple,
    "Late" = RSC_NOL_Late_purple
  )) +
  
  scale_fill_manual(values = c(
   "Early" = RSC_NOL_Early_purple,
    "Late" = RSC_NOL_Late_purple
  )) +
  
  # Line types
  scale_linetype_manual(values = c(
    "Early" = "dashed",
    "Late" = "solid"
  )) +
  
   # Y-axis formatting ONLY
  scale_y_continuous(
    breaks = seq(0, 0.12, by = 0.02),
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
  ) +
  
  # Zoom display WITHOUT removing data
  coord_cartesian(
     ylim = c(0, 0.12)
  ) + 
  
   # Labels
  labs(x = "Hall",
       y = "Representational similarity (r)",
       title = "(1) NOL Learning Stage RSC",
       color = "Learning stage",
       fill = "Learning stage",
       linetype = "Learning stage") +
  
  
  # Clean theme
  theme_classic(base_size = 14)

print(p_NOL_RSC)

# Save graph
ggsave(
  "NOL_Early-Late_RSC.tiff",
  plot = p_NOL_RSC,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)


```

**************************
RSComplex OL
**************************

-> Repeated meausures ANOVA Learning stage (Early,Late) x Hall (Cue period, 1st hall, 2nd hall, goal)
(1) OL RSComplex 
```{r}

# 1) Filtering + Subject means
plot_data_OL_RSComplex <- long_parsed %>%
  filter(
    ROI %in% c("Complex"),     # ✅ correct ROI name in data
    str_detect(raw_name, "ol_new_early_w_ol_new_early_|ol_new_late_w_ol_new_late_") 
  ) %>%
  group_by(ID, learningStage, Hall) %>%
  summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop")

stopifnot(nrow(plot_data_OL_RSComplex) > 0)

print(plot_data_OL_RSComplex)

```
ANOVA
```{r}


anova_model_OL_RSComplex <- aov_ez(
  id = "ID",
  dv = "RSA",
  data = plot_data_OL_RSComplex,
  within = c("learningStage", "Hall"),
  type = 3,
  anova_table = list(
    correction = "GG",
    es = "pes"
  )
)


# Extract ANOVA table as data frame
anova_table_OL_RSComplex <- as.data.frame(anova_model_OL_RSComplex$anova_table) |>
  tibble::rownames_to_column(var = "Effect")


 
# Format p-values like JASP
format_p <- function(p) {
  if (p < .001) "< .001"
  else paste0("= ", format(round(p, 3), nsmall = 3))
}


# Clean ANOVA table
 
niceAnovaTable_OL_RSComplex <- data.frame(
  Effect = anova_table_OL_RSComplex$Effect,
  df1    = anova_table_OL_RSComplex[[1]],
  df2    = anova_table_OL_RSComplex[[2]],
  F      = round(anova_table_OL_RSComplex$F, 3),
  p      = sapply(anova_table_OL_RSComplex$`Pr(>F)`, format_p),
  `η²p`  = round(anova_table_OL_RSComplex$pes, 3)
)

print(niceAnovaTable_OL_RSComplex)

```
POST-HOCs 
```{r}
emm_Hall_OL_RSComplex <- emmeans(anova_model_OL_RSComplex, ~ Hall)

posthoc_OL_RSComplex <- pairs(emm_Hall_OL_RSComplex, adjust = "holm")

holm_table_OL_RSComplex <- summary(posthoc_OL_RSComplex) |>
  as.data.frame() |>
  mutate(
    estimate = round(estimate, 3),
    SE       = round(SE, 3),
    t        = round(t.ratio, 2),

    cohen_d  = round(abs(t.ratio) / sqrt(df + 1), 3),

    pHolms = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ paste0("= ", sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3))))
    ),

    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE ~ ""
    )
  )

print(holm_table_OL_RSComplex)
```
DESCRIPTIVES

```{r}

descriptives_OL_RSComplex <- plot_data_OL_RSComplex %>%
  group_by(learningStage, Hall) %>%
  summarise(
    N = n(),
    Mean = mean(RSA, na.rm = TRUE),
    SD = sd(RSA, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(across(c(Mean, SD, SE), ~ round(.x, 3)))

print(descriptives_OL_RSComplex)

```
PLOTTING OL RSComplex (LearningStage x Hall)

```{r}

# Plot with ribbon for mean ± SE
p_OL_RSComplex <- ggplot(plot_data_OL_RSComplex,
       aes(x = Hall,
           y = RSA,
           color = learningStage,
           fill = learningStage,
           group = learningStage)) +
  
# Ribbon for standard error (mean ± SE)
  stat_summary(fun.data = mean_se,
               geom = "ribbon",
               alpha = 0.15,
               color = NA) +
  
# Mean line with Early dashed, Late solid
  stat_summary(fun = mean,
               geom = "line",
               aes(linetype = learningStage),
               linewidth = 1) +
  
# Mean points
  stat_summary(fun = mean,
               geom = "point",
               size = 3) +
  
# Custom colors
  
  scale_color_manual(values = c(
    "Early" = Complex_OL_Early_blue,
    "Late" = Complex_OL_Late_blue
  )) +
  
  scale_fill_manual(values = c(
   "Early" = Complex_OL_Early_blue,
    "Late" = Complex_OL_Late_blue
  )) +
  
# Line types
  scale_linetype_manual(values = c(
    "Early" = "dashed",
    "Late" = "solid"
  )) +
  
  # Y-axis formatting ONLY
  scale_y_continuous(
    breaks = seq(0, 0.12, by = 0.02),
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
  ) +
  
  # Zoom display WITHOUT removing data
  coord_cartesian(
     ylim = c(0, 0.12)
  ) + 
    
  # Labels
  labs(x = "Hall",
       y = "Representational similarity (r)",
       title = "(1) OL Learning Stage RSComplex",
       color = "Learning stage",
       fill = "Learning stage",
       linetype = "Learning stage") +
  
  
# Clean theme
  theme_classic(base_size = 14)

print(p_OL_RSComplex)

# Save graph
ggsave(
  "OL_Early-Late_RSComplex.tiff",
  plot = p_OL_RSComplex,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)
```

**************************
RSComplex NOL
**************************

-> Repeated meausures ANOVA Learning stage (Early,Late) x Hall (Cue period, 1st hall, 2nd hall, goal)
(1) NOL RSComplex 
```{r}

# 1) Filtering + Subject means
plot_data_NOL_RSComplex <- long_parsed %>%
  filter(
    ROI %in% c("Complex"),     # ✅ correct ROI name in data
    str_detect(raw_name, "nol_new_early_w_nol_new_early_|nol_new_late_w_nol_new_late_") 
  ) %>%
  group_by(ID, learningStage, Hall) %>%
  summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop")

stopifnot(nrow(plot_data_NOL_RSComplex) > 0)

print(plot_data_NOL_RSComplex)

```
ANOVA
```{r}

anova_model_NOL_RSComplex <- aov_ez(
  id = "ID",
  dv = "RSA",
  data = plot_data_NOL_RSComplex,
  within = c("learningStage", "Hall"),
  type = 3,
  anova_table = list(
    correction = "GG",
    es = "pes"
  )
)


# Extract ANOVA table as data frame
anova_table_NOL_RSComplex <- as.data.frame(anova_model_NOL_RSComplex$anova_table) |>
  tibble::rownames_to_column(var = "Effect")


 
# Format p-values like JASP
format_p <- function(p) {
  if (p < .001) "< .001"
  else paste0("= ", format(round(p, 3), nsmall = 3))
}


# Clean ANOVA table
 
niceAnovaTable_NOL_RSComplex <- data.frame(
  Effect = anova_table_NOL_RSComplex$Effect,
  df1    = anova_table_NOL_RSComplex[[1]],
  df2    = anova_table_NOL_RSComplex[[2]],
  F      = round(anova_table_NOL_RSComplex$F, 3),
  p      = sapply(anova_table_NOL_RSComplex$`Pr(>F)`, format_p),
  `η²p`  = round(anova_table_NOL_RSComplex$pes, 3)
)

print(niceAnovaTable_NOL_RSComplex)

```
POST-HOCs 
```{r}
# 4 main effect of HAll
emm_Hall_NOL_RSComplex <- emmeans(anova_model_NOL_RSComplex, ~ Hall)

posthoc_NOL_RSComplex <- pairs(emm_Hall_NOL_RSComplex, adjust = "holm")

holm_table_NOL_RSComplex <- summary(posthoc_NOL_RSComplex) |>
  as.data.frame() |>
  mutate(
    estimate = round(estimate, 3),
    SE       = round(SE, 3),
    t        = round(t.ratio, 2),

    cohen_d  = round(abs(t.ratio) / sqrt(df + 1), 3),

    pHolms = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ paste0("= ", sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3))))
    ),

    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE ~ ""
    )
  )

print(holm_table_NOL_RSComplex)

# # If significant interaction (Note no longer significant with GG sphericity correction)
# emm_interaction <- emmeans(anova_model_NOL_RSComplex, ~ learningStage * Hall)
# 
# # Holm correction
# posthoc_holm_inter <- pairs(emm_interaction, adjust = "holm")
# 
# holm_table_NOL_RSComplex_Inter <- summary(posthoc_holm_inter) |>
#   as.data.frame() |>
#   mutate(
#     estimate = round(estimate, 3),
#     SE = round(SE, 3),
#     cohen_d = round(estimate / (SE * sqrt(2)), 3),  
#     t = round(t.ratio, 2),
#     pHolms = ifelse(p.value < .001,
#                "< 0.001",
#                paste0("= ", round(p.value, 3)))
#   )
# 
# print(holm_table_NOL_RSComplex_Inter)

```
DESCRIPTIVES

```{r}

descriptives_NOL_RSComplex <- plot_data_NOL_RSComplex %>%
  group_by(learningStage, Hall) %>%
  summarise(
    N = n(),
    Mean = mean(RSA, na.rm = TRUE),
    SD = sd(RSA, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(across(c(Mean, SD, SE), ~ round(.x, 3)))

print(descriptives_NOL_RSComplex)
```
PLOTTING NOL RSComplex (LearningStage x Hall)
********** <3 ********************

```{r}

# 2) Plot with ribbon for mean ± SE
p_NOL_RSComplex <- ggplot(plot_data_NOL_RSComplex,
       aes(x = Hall,
           y = RSA,
           color = learningStage,
           fill = learningStage,
           group = learningStage)) +
  
# Ribbon for standard error
  stat_summary(fun.data = mean_se,
               geom = "ribbon",
               alpha = 0.15,
               color = NA) +
  
# Mean line with Early dashed, Late solid
  stat_summary(fun = mean,
               geom = "line",
               aes(linetype = learningStage),
               linewidth = 1) +
  
# Mean points
  stat_summary(fun = mean,
               geom = "point",
               size = 3) +
  
# Custom colors
  
  scale_color_manual(values = c(
    "Early" = Complex_NOL_Early_blue,
    "Late" = Complex_NOL_Late_blue
  )) +
  
  scale_fill_manual(values = c(
   "Early" = Complex_NOL_Early_blue,
    "Late" = Complex_NOL_Late_blue
  )) +
  
# Line types
  scale_linetype_manual(values = c(
    "Early" = "dashed",
    "Late" = "solid"
  )) +
  
 # Y-axis formatting ONLY
  scale_y_continuous(
    breaks = seq(0, 0.12, by = 0.02),
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
  ) +
  
  # Zoom display WITHOUT removing data
  coord_cartesian(
     ylim = c(0, 0.12)
  ) + 
  
  # Labels
  labs(x = "Hall",
       y = "Representational similarity (r)",
       title = "(1) NOL Learning Stage RSComplex",
       color = "Learning stage",
       fill = "Learning stage",
       linetype = "Learning stage") +
  
  
# Clean theme
  theme_classic(base_size = 14)

print(p_NOL_RSComplex)

# Save graph
ggsave(
  "NOL_Early-Late_RSComplex.tiff",
  plot = p_NOL_RSComplex,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)

```
*****************************************
2) NATURE OF REPRESENTATIONS 
**************************

   2a) 3 ROIs: HC vs RSC vs RSComplex 
**************************
#   OL EARLY 
```{r}
#  Repeated measures ANOVA
# 1) Filtering + Subject means
plot_data_OL_Early_HCvsRSCvsComplex <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "RSC", "Hippocampus"),  # ROI filter
    str_detect(raw_name, "ol_new_early_w_ol_new_early_")  
  ) %>%
  group_by(ID, ROI, Hall) %>%
  summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop")

# ✅ FAIL FAST HERE
stopifnot(nrow(plot_data_OL_Early_HCvsRSCvsComplex) > 0)

print(plot_data_OL_Early_HCvsRSCvsComplex)

```
ANOVA
```{r}

 # ANOVA

anova_model_OL_Early_HCvsRSCvsComplex <- aov_ez(
  id = "ID",
  dv = "RSA",
  data = plot_data_OL_Early_HCvsRSCvsComplex,
  within = c("ROI", "Hall"),
  type = 3,  # Type III SS (JASP style)
  anova_table = list(
    correction = "GG",  # Greenhouse-Geisser correction for Hall (4 levels)
    es = "pes"          # partial eta squared
  )
)

# Extract ANOVA table as data frame
anova_table_OL_Early_HCvsRSCvsComplex <- as.data.frame(anova_model_OL_Early_HCvsRSCvsComplex$anova_table) |>
  tibble::rownames_to_column(var = "Effect")


# Format p-values like JASP
format_p <- function(p) {
  if (p < .001) "< .001"
  else paste0("= ", format(round(p, 3), nsmall = 3))
}


# Clean ANOVA table
 
niceAnovaTable_OL_Early_HCvsRSCvsComplex <- data.frame(
  Effect = anova_table_OL_Early_HCvsRSCvsComplex$Effect,
  df1    = anova_table_OL_Early_HCvsRSCvsComplex[[1]],
  df2    = anova_table_OL_Early_HCvsRSCvsComplex[[2]],
  F      = round(anova_table_OL_Early_HCvsRSCvsComplex$F, 3),
  p      = sapply(anova_table_OL_Early_HCvsRSCvsComplex$`Pr(>F)`, format_p),
  `η²p`  = round(anova_table_OL_Early_HCvsRSCvsComplex$pes, 3)
)

print(niceAnovaTable_OL_Early_HCvsRSCvsComplex)

```
* POST-HOCS
```{r}

# full (cross-ROI comparisons) ROI × Hall interaction ( = JASP)
emm_interaction_OL_Early_HCvsRSCvsComplex <- emmeans(
  anova_model_OL_Early_HCvsRSCvsComplex,
  ~ Hall * ROI
)

posthoc_interaction_OL_Early_HCvsRSCvsComplex <- pairs(
  emm_interaction_OL_Early_HCvsRSCvsComplex,
  adjust = "holm"
)


# Interaction table 
interaction_table_OL_Early_HCvsRSCvsComplex <-
  summary(posthoc_interaction_OL_Early_HCvsRSCvsComplex) |>
  as.data.frame() |>
  mutate(
    estimate = round(estimate, 3),
    SE       = round(SE, 3),
    t        = round(t.ratio, 2),


    pHolm = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ paste0(
        "= ",
        sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3)))
      )
    ),

    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE           ~ ""
    )
  ) |>
  select(
    contrast,
    estimate,
    SE,
    df,
    t,
    pHolm,
    sig
  )

print(interaction_table_OL_Early_HCvsRSCvsComplex)


# # If significant interaction (Note no longer significant with GG sphericity correction)
# emm_interaction <- emmeans(anova_model_NOL_RSComplex, ~ learningStage * Hall)
# 
# # Holm correction
# posthoc_holm_inter <- pairs(emm_interaction, adjust = "holm")
# 
# holm_table_NOL_RSComplex_Inter <- summary(posthoc_holm_inter) |>
#   as.data.frame() |>
#   mutate(
#     estimate = round(estimate, 3),
#     SE = round(SE, 3),
#     cohen_d = round(estimate / (SE * sqrt(2)), 3),  
#     t = round(t.ratio, 2),
#     pHolms = ifelse(p.value < .001,
#                "< 0.001",
#                paste0("= ", round(p.value, 3)))
#   )
# 
# print(holm_table_NOL_RSComplex_Inter)

```

DESCRIPTIVES
```{r}
###) Descriptives
descriptives_OL_Early_HCvsRSCvsComplex <- plot_data_OL_Early_HCvsRSCvsComplex %>%
  group_by(ROI, Hall) %>%
  summarise(
    N = n(),
    Mean = mean(RSA, na.rm = TRUE),
    SD = sd(RSA, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(across(c(Mean, SD, SE), ~ round(.x, 3)))

print(descriptives_OL_Early_HCvsRSCvsComplex)
```
# PLOTTING OL EARLY HC vs RSC vs RSComplex
```{r}
# 1) Make sure ROI is a factor following the order I want and not the by-default alphabetical one
#plot_data$ROI <- factor(plot_data$ROI, levels = c("Hippocampus", "Complex"))
# 2) Plot with ribbon for mean ± SE
p_OL_Early_HCvsRSCvsComplex <- ggplot(plot_data_OL_Early_HCvsRSCvsComplex,
       aes(x = Hall,
           y = RSA,
           color = ROI,
           fill = ROI,
           group = ROI)) +

  # Ribbon for standard error
  stat_summary(fun.data = mean_se,
               geom = "ribbon",
               alpha = 0.15,
               color = NA) +

  # Mean line
  stat_summary(fun = mean,
               geom = "line",
               linewidth = 1) +

  # Mean points
  stat_summary(fun = mean,
               geom = "point",
               size = 3) +

  # Custom colors

  scale_color_manual(values = c(
    "Hippocampus" = HC_OL_Early_orange,
     "RSC" = RSC_OL_Early_purple,
    "Complex" = Complex_OL_Early_blue
  ),

   labels = c("Hippocampus", "RSC", "RSComplex")) +

  scale_fill_manual(values = c(
    "Hippocampus" = HC_OL_Early_orange,
    "RSC" = RSC_OL_Early_purple,
    "Complex" = Complex_OL_Early_blue
  ),

  labels = c("Hippocampus", "RSC", "RSComplex")) +

  # Labels
  labs(x = "Hall",
       y = "Representational similarity (r)",
       title = "(2a) OL EARLY HC RSC vs RSComplex",
       #subtitle = "(2a) OL EARLY HCvsComplex",
       color = "ROI",
       fill = "ROI") +
  
  # Y-axis formatting
  scale_y_continuous(
    breaks = seq(0, 0.10, by = 0.02),
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
    ) +
  
  # Zoom display WITHOUT removing data
  coord_cartesian(
    ylim = c(0, 0.107)
  ) +

  # Clean theme
  theme_classic(base_size = 14)

print (p_OL_Early_HCvsRSCvsComplex)

# Save graph
ggsave(
  "OL_Early_HCvsRSCvsComplex.tiff",
  plot = p_OL_Early_HCvsRSCvsComplex,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)


```
 2a) HC vs RSC vs RSComplex
 
#    OL LATE

```{r}
#  Repeated measures ANOVA
# 1) Filtering + Subject means
plot_data_OL_Late_HCvsRSCvsComplex <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "RSC", "Hippocampus"),     
    str_detect(raw_name, "ol_new_late_w_ol_new_late_")  
  ) %>%
  group_by(ID, ROI, Hall) %>%
  summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop")

# ✅ FAIL FAST HERE
stopifnot(nrow(plot_data_OL_Late_HCvsRSCvsComplex) > 0)

print(plot_data_OL_Late_HCvsRSCvsComplex)

```
ANOVA
```{r}

  # ANOVA

anova_model_OL_Late_HCvsRSCvsComplex <- aov_ez(
  id = "ID",
  dv = "RSA",
  data = plot_data_OL_Late_HCvsRSCvsComplex,
  within = c("ROI", "Hall"),
  type = 3,  # Type III SS (JASP style)
  anova_table = list(
    correction = "GG",  # Greenhouse-Geisser correction for Hall (4 levels)
    es = "pes"          # partial eta squared
  )
)

# Extract ANOVA table as data frame
anova_table_OL_Late_HCvsRSCvsComplex <- as.data.frame(anova_model_OL_Late_HCvsRSCvsComplex$anova_table) |>
  tibble::rownames_to_column(var = "Effect")


# Format p-values like JASP
format_p <- function(p) {
  if (p < .001) "< .001"
  else paste0("= ", format(round(p, 3), nsmall = 3))
}


# Clean ANOVA table
 
niceAnovaTable_OL_Late_HCvsRSCvsComplex <- data.frame(
  Effect = anova_table_OL_Late_HCvsRSCvsComplex$Effect,
  df1    = anova_table_OL_Late_HCvsRSCvsComplex[[1]],
  df2    = anova_table_OL_Late_HCvsRSCvsComplex[[2]],
  F      = round(anova_table_OL_Late_HCvsRSCvsComplex$F, 3),
  p      = sapply(anova_table_OL_Late_HCvsRSCvsComplex$`Pr(>F)`, format_p),
  `η²p`  = round(anova_table_OL_Late_HCvsRSCvsComplex$pes, 3)
)

print(niceAnovaTable_OL_Late_HCvsRSCvsComplex)


```
* POST-HOCS
```{r}

# full (cross-ROI comparisons) ROI × Hall interaction ( = JASP)
emm_interaction_OL_Late_HCvsRSCvsComplex <- emmeans(
  anova_model_OL_Late_HCvsRSCvsComplex,
  ~ Hall * ROI
)

posthoc_interaction_OL_Late_HCvsRSCvsComplex <- pairs(
  emm_interaction_OL_Late_HCvsRSCvsComplex,
  adjust = "holm"
)


# Interaction table 
interaction_table_OL_Late_HCvsRSCvsComplex <-
  summary(posthoc_interaction_OL_Late_HCvsRSCvsComplex) |>
  as.data.frame() |>
  mutate(
    estimate = round(estimate, 3),
    SE       = round(SE, 3),
    t        = round(t.ratio, 2),


    pHolm = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ paste0(
        "= ",
        sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3)))
      )
    ),

    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE           ~ ""
    )
  ) |>
  select(
    contrast,
    estimate,
    SE,
    df,
    t,
    pHolm,
    sig
  )

print(interaction_table_OL_Late_HCvsRSCvsComplex)




# # If significant interaction (Note no longer significant with GG sphericity correction)
# emm_interaction <- emmeans(anova_model_NOL_RSComplex, ~ learningStage * Hall)
# 
# # Holm correction
# posthoc_holm_inter <- pairs(emm_interaction, adjust = "holm")
# 
# holm_table_NOL_RSComplex_Inter <- summary(posthoc_holm_inter) |>
#   as.data.frame() |>
#   mutate(
#     estimate = round(estimate, 3),
#     SE = round(SE, 3),
#     cohen_d = round(estimate / (SE * sqrt(2)), 3),  
#     t = round(t.ratio, 2),
#     pHolms = ifelse(p.value < .001,
#                "< 0.001",
#                paste0("= ", round(p.value, 3)))
#   )
# 
# print(holm_table_NOL_RSComplex_Inter)

```

DESCRIPTIVES
```{r}

descriptives_OL_Late_HCvsRSCvsComplex <- plot_data_OL_Late_HCvsRSCvsComplex %>%
  group_by(ROI, Hall) %>%
  summarise(
    N = n(),
    Mean = mean(RSA, na.rm = TRUE),
    SD = sd(RSA, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(across(c(Mean, SD, SE), ~ round(.x, 3)))

print(descriptives_OL_Late_HCvsRSCvsComplex)
```
# PLOTTING OL LATE HC vs RSC vs RSComplex
```{r}
# 1) Make sure ROI is a factor following the order I want and not the by-default alphabetical one
#plot_data$ROI <- factor(plot_data$ROI, levels = c("Hippocampus", "Complex"))
# 2) Plot with ribbon for mean ± SE
p_OL_Late_HCvsRSCvsComplex <- ggplot(plot_data_OL_Late_HCvsRSCvsComplex,
       aes(x = Hall,
           y = RSA,
           color = ROI,
           fill = ROI,
           group = ROI)) +

  # Ribbon for standard error
  stat_summary(fun.data = mean_se,
               geom = "ribbon",
               alpha = 0.15,
               color = NA) +

  # Mean line
  stat_summary(fun = mean,
               geom = "line",
               linewidth = 1) +

  # Mean points
  stat_summary(fun = mean,
               geom = "point",
               size = 3) +

  # Custom colors

  scale_color_manual(values = c(
    "Hippocampus" = HC_OL_Late_orange,
    "RSC" = RSC_OL_Late_purple,
    "Complex" = Complex_OL_Late_blue
  ),

   labels = c("Hippocampus", "RSC","RSComplex")) +

  scale_fill_manual(values = c(
    "Hippocampus" = HC_OL_Late_orange,
    "RSC" = RSC_OL_Late_purple,
    "Complex" = Complex_OL_Late_blue
  ),

  labels = c("Hippocampus", "RSC","RSComplex")) +

  # Labels
  labs(x = "Hall",
       y = "Representational similarity (r)",
       title = "(2a) OL LATE HC vs RSC vs RSComplex",
       #subtitle = "(2a) OL LATE HCvsComplex",
       color = "ROI",
       fill = "ROI") +
  
  # Y-axis formatting
  scale_y_continuous(
    breaks = seq(0, 0.10, by = 0.02),
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
    ) +
  
  # Zoom display WITHOUT removing data
  coord_cartesian(
    ylim = c(0, 0.107)
  ) +

  # Clean theme
  theme_classic(base_size = 14)

print(p_OL_Late_HCvsRSCvsComplex)

# Save graph
ggsave(
  "OL_Late_HCvsRSCvsComplex.tiff",
  plot = p_OL_Late_HCvsRSCvsComplex,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)


```
********************************
2a) HC vs RSC vs RSComplex

#  NOL Early 

```{r}

# 1) Filtering + Subject means
plot_data_NOL_Early_HCvsRSCvsComplex <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "RSC", "Hippocampus"),     
    str_detect(raw_name, "nol_new_early_w_nol_new_early_")  
  ) %>%
  group_by(ID, ROI, Hall) %>%
  summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop")
  

# ✅ FAIL FAST HERE
stopifnot(nrow(plot_data_NOL_Early_HCvsRSCvsComplex) > 0)

print(plot_data_NOL_Early_HCvsRSCvsComplex)

```
ANOVA
```{r}

 # ANOVA
# summary
# options(contrasts = c("contr.sum", "contr.poly")) #same analysis as in JASP
anova_model_NOL_Early_HCvsRSCvsComplex <- aov_ez(
  id = "ID",
  dv = "RSA",
  data = plot_data_NOL_Early_HCvsRSCvsComplex,
  within = c("ROI", "Hall"),
  type = 3, # Type III SS (JASP style)
  anova_table = list(
  correction = "GG",
    es = "pes"   # partial eta squared
  )
)

# ANOVA TABLE
anova_table_NOL_Early_HCvsRSCvsComplex <- as.data.frame(anova_model_NOL_Early_HCvsRSCvsComplex$anova_table) |>
  tibble::rownames_to_column(var = "Effect")

# Format p-values like JASP
format_p <- function(p) {
  if (p < .001) {
    "< .001"
  } else {
    paste0("= ", format(round(p, 3), nsmall = 3))
  }
}

# Clean ANOVA table
niceAnovaTable_NOL_Early_HCvsRSCvsComplex <- data.frame(
  Effect = anova_table_NOL_Early_HCvsRSCvsComplex$Effect,
  df1    = anova_table_NOL_Early_HCvsRSCvsComplex[[1]],
  df2    = anova_table_NOL_Early_HCvsRSCvsComplex[[2]],
  F      = round(anova_table_NOL_Early_HCvsRSCvsComplex$F, 3),
  p      = sapply(anova_table_NOL_Early_HCvsRSCvsComplex$`Pr(>F)`, format_p),
  `η²p`  = round(anova_table_NOL_Early_HCvsRSCvsComplex$pes, 3)
)

print(niceAnovaTable_NOL_Early_HCvsRSCvsComplex)


```
* POST-HOCS
*** Sig interaction ***
```{r}
 # full (cross-ROI comparisons) ROI × Hall interaction ( = JASP)
emm_interaction_NOL_Early_HCvsRSCvsComplex <- emmeans(
  anova_model_NOL_Early_HCvsRSCvsComplex,
  ~ Hall * ROI
)

posthoc_interaction_NOL_Early_HCvsRSCvsComplex <- pairs(
  emm_interaction_NOL_Early_HCvsRSCvsComplex,
  adjust = "holm"
)


# Interaction table 
interaction_table_NOL_Early_HCvsRSCvsComplex <-
  summary(posthoc_interaction_NOL_Early_HCvsRSCvsComplex) |>
  as.data.frame() |>
  mutate(
    estimate = round(estimate, 3),
    SE       = round(SE, 3),
    t        = round(t.ratio, 2),


    pHolm = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ paste0(
        "= ",
        sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3)))
      )
    ),

    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE           ~ ""
    )
  ) |>
  select(
    contrast,
    estimate,
    SE,
    df,
    t,
    pHolm,
    sig
  )

print(interaction_table_NOL_Early_HCvsRSCvsComplex)

```
DESCRIPTIVES
```{r}
###) Descriptives
descriptives_NOL_Early_HCvsRSCvsComplex <- plot_data_NOL_Early_HCvsRSCvsComplex %>%
  group_by(ROI, Hall) %>%
  summarise(
    N = n(),
    Mean = mean(RSA, na.rm = TRUE),
    SD = sd(RSA, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(across(c(Mean, SD, SE), ~ round(.x, 3)))

print(descriptives_NOL_Early_HCvsRSCvsComplex)
```
PLOTTING NOL Early HC vs RSC vs RSComplex
```{r}

# Plot with ribbon for mean ± SE
p_NOL_Early_HCvsRSCvsComplex <- ggplot(plot_data_NOL_Early_HCvsRSCvsComplex,
       aes(x = Hall,
           y = RSA,
           color = ROI,
           fill = ROI,
           group = ROI)) +

  # Ribbon for standard error
  stat_summary(fun.data = mean_se,
               geom = "ribbon",
               alpha = 0.15,
               color = NA) +

  # Mean line
  stat_summary(fun = mean,
               geom = "line",
               linewidth = 1) +

  # Mean points
  stat_summary(fun = mean,
               geom = "point",
               size = 3) +

  # Custom colors

  scale_color_manual(values = c(
    "Hippocampus" = HC_NOL_Early_orange,
     "RSC" = RSC_NOL_Early_purple,
    "Complex" = Complex_NOL_Early_blue

  ),
  labels = c("Hippocampus","RSC" ,"RSComplex")) +

  scale_fill_manual(values = c(
    "Hippocampus" = HC_NOL_Early_orange,
     "RSC" = RSC_NOL_Early_purple,
    "Complex" = Complex_NOL_Early_blue

  ),

  labels = c("Hippocampus", "RSC", "RSComplex")) +

  # Labels
  labs(x = "Hall",
       y = "Representational similarity (r)",
       title = "(2a) NOL EARLY HC vs RSC vs RSComplex",
       #subtitle = "(2a) OL EARLY HCvsComplex",
       color = "ROI",
       fill = "ROI") +

  # Y-axis formatting
  scale_y_continuous(
    breaks = seq(0, 0.10, by = 0.02),
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
    ) +
  
  # Zoom display WITHOUT removing data
  coord_cartesian(
    ylim = c(0, 0.107)
  ) +
  
  # Clean theme
  theme_classic(base_size = 14)

print(p_NOL_Early_HCvsRSCvsComplex)

# Save graph
ggsave(
  "NOL_Early_HCvsRSCvsComplex.tiff",
  plot = p_NOL_Early_HCvsRSCvsComplex,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)



```
2a) HC vs RSC vs Complex 

#  NOL LATE
```{r}
# (2a) NOL LATE HC vs Complex 

# 1) Filtering + Subject means
plot_data_NOL_Late_HCvsRSCvsComplex <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "RSC","Hippocampus"),     # ROI filter
    str_detect(raw_name, "nol_new_late_w_nol_new_late_")  
  ) %>%
  group_by(ID, ROI, Hall) %>%
  summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop")
  
# ✅ FAIL FAST HERE
stopifnot(nrow(plot_data_NOL_Late_HCvsRSCvsComplex) > 0)

print(plot_data_NOL_Late_HCvsRSCvsComplex)

```
ANOVA
```{r}

 # ANOVA
anova_model_NOL_Late_HCvsRSCvsComplex <- aov_ez(
  id = "ID",
  dv = "RSA",
  data = plot_data_NOL_Late_HCvsRSCvsComplex,
  within = c("ROI", "Hall"),
  type = 3, # Type III SS (JASP style)
  anova_table = list(
  correction = "GG",
    es = "pes"   # partial eta squared
  )
)

# ANOVA TABLE
anova_table_NOL_Late_HCvsRSCvsComplex <- as.data.frame(anova_model_NOL_Late_HCvsRSCvsComplex$anova_table) |>
  tibble::rownames_to_column(var = "Effect")

# Format p-values like JASP
format_p <- function(p) {
  if (p < .001) {
    "< .001"
  } else {
    paste0("= ", format(round(p, 3), nsmall = 3))
  }
}

# Clean ANOVA table
niceAnovaTable_NOL_Late_HCvsRSCvsComplex <- data.frame(
  Effect = anova_table_NOL_Late_HCvsRSCvsComplex$Effect,
  df1    = anova_table_NOL_Late_HCvsRSCvsComplex[[1]],
  df2    = anova_table_NOL_Late_HCvsRSCvsComplex[[2]],
  F      = round(anova_table_NOL_Late_HCvsRSCvsComplex$F, 3),
  p      = sapply(anova_table_NOL_Late_HCvsRSCvsComplex$`Pr(>F)`, format_p),
  `η²p`  = round(anova_table_NOL_Late_HCvsRSCvsComplex$pes, 3)
)

print(niceAnovaTable_NOL_Late_HCvsRSCvsComplex)

```
* POST-HOCS
```{r}
 # full (cross-ROI comparisons) ROI × Hall interaction ( = JASP)
emm_interaction_NOL_Late_HCvsRSCvsComplex <- emmeans(
  anova_model_NOL_Late_HCvsRSCvsComplex,
  ~ Hall * ROI
)

posthoc_interaction_NOL_Late_HCvsRSCvsComplex <- pairs(
  emm_interaction_NOL_Late_HCvsRSCvsComplex,
  adjust = "holm"
)


# Interaction table 
interaction_table_NOL_Late_HCvsRSCvsComplex <-
  summary(posthoc_interaction_NOL_Late_HCvsRSCvsComplex) |>
  as.data.frame() |>
  mutate(
    estimate = round(estimate, 3),
    SE       = round(SE, 3),
    t        = round(t.ratio, 2),


    pHolm = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ paste0(
        "= ",
        sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3)))
      )
    ),

    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      TRUE           ~ ""
    )
  ) |>
  select(
    contrast,
    estimate,
    SE,
    df,
    t,
    pHolm,
    sig
  )

print(interaction_table_NOL_Late_HCvsRSCvsComplex)
# NO INTERACTION

# *** NO INTERACTION ***

# Simple effects of Hall within ROI: Averaging across ROI, do halls differ?
# 4 main effects of HAll (when interaction is not significant)
# emm_Hall_NOL_Early_HCvsComplex <- emmeans(anova_model_NOL_Late_HCvsComplex, ~ Hall)
# 
# posthoc_Hall_NOL_Late_HCvsComplex <- pairs(emm_Hall_NOL_Late_HCvsComplex, adjust = "holm")
# 
# holm_table_NOL_Late_HCvsComplex <- summary(posthoc_Hall_NOL_Late_HCvsComplex) |>
#   as.data.frame() |>
#   mutate(
#     estimate = round(estimate, 3),
#     SE       = round(SE, 3),
#     t        = round(t.ratio, 2),
# 
#     cohen_d  = round(abs(t.ratio) / sqrt(df + 1), 3),
# 
#     pHolms = case_when(
#       p.value < .001 ~ "< .001",
#       TRUE ~ paste0("= ", sub("^0\\.", ".", sprintf("%.3f", round(p.value, 3))))
#     ),
# 
#     sig = case_when(
#       p.value < .001 ~ "***",
#       p.value < .01  ~ "**",
#       p.value < .05  ~ "*",
#       TRUE ~ ""
#     )
#   )
# 
# print(holm_table_NOL_Late_HCvsComplex)
```

DESCRIPTIVES
```{r}
###) Descriptives
descriptives_NOL_Late_HCvsRSCvsComplex <- plot_data_NOL_Late_HCvsRSCvsComplex %>%
  group_by(ROI, Hall) %>%
  summarise(
    N = n(),
    Mean = mean(RSA, na.rm = TRUE),
    SD = sd(RSA, na.rm = TRUE),
    SE = SD / sqrt(N),
    .groups = "drop"
  ) %>%
  mutate(across(c(Mean, SD, SE), ~ round(.x, 3)))

print(descriptives_NOL_Late_HCvsRSCvsComplex)
```
# PLOTTING NOL LATE HC vs RSC vs Complex
```{r}

# Plot with ribbon for mean ± SE
p_NOL_Late_HCvsRSCvsComplex <- ggplot(plot_data_NOL_Late_HCvsRSCvsComplex,
       aes(x = Hall,
           y = RSA,
           color = ROI,
           fill = ROI,
           group = ROI)) +

  # Ribbon for standard error
  stat_summary(fun.data = mean_se,
               geom = "ribbon",
               alpha = 0.15,
               color = NA) +

  # Mean line
  stat_summary(fun = mean,
               geom = "line",
               linewidth = 1) +

  # Mean points
  stat_summary(fun = mean,
               geom = "point",
               size = 3) +

  # Custom colors

  scale_color_manual(values = c(
    "Hippocampus" = HC_NOL_Late_orange,
     "RSC" = RSC_NOL_Late_purple,
    "Complex" = Complex_NOL_Late_blue

  ),
  
  labels = c("Hippocampus","RSC" ,"RSComplex")) +

  scale_fill_manual(values = c(
    "Hippocampus" = HC_NOL_Late_orange,
     "RSC" = RSC_NOL_Late_purple,
    "Complex" = Complex_NOL_Late_blue

  ),

  labels = c("Hippocampus", "RSC", "RSComplex")) +

  # Labels
  labs(x = "Hall",
       y = "Representational similarity (r)",
       title = "(2a) NOL LATE HC vs RSC vs RSComplex",
       #subtitle = "(2a) OL EARLY HCvsComplex",
       color = "ROI",
       fill = "ROI") +
  
  # Y-axis formatting
  scale_y_continuous(
    breaks = seq(0, 0.10, by = 0.02),
    labels = scales::label_number(accuracy = 0.01),
    expand = c(0, 0)
    ) +
  
  # Zoom display WITHOUT removing data
  coord_cartesian(
    ylim = c(0, 0.107)
  ) +

  # Clean theme
  theme_classic(base_size = 14)

print(p_NOL_Late_HCvsRSCvsComplex)

# Save graph
ggsave(
  "NOL_Late_HCvsRSCvsComplex.tiff",
  plot = p_NOL_Late_HCvsRSCvsComplex,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)


```

****************************
2) NATURE OF REPRESENTATIONS
****************************

(2b) REPULSION, SEPARATION OR COMPLETION REPRESENTATIONS?
******************************************************
#  2b.1) Hippocampus (EARLY+LATE)

```{r}

# 1) Filter data
## Early
early_data_nature_HC <- long_parsed %>%
  filter(
    Overlappicity == "OL",
    learningStage == "Early",
    ROI == "Hippocampus",
    str_detect(
      raw_name,
      "ol_new_early_w_ol_new_early|ol_new_early_w_ol_old_early"
    )
  )

# Fail fast (defensive)
stopifnot(nrow(early_data_nature_HC) > 0)

## Late
late_data_nature_HC <- long_parsed %>%
  filter(
    Overlappicity == "OL",
    learningStage == "Late",
    ROI == "Hippocampus",
    str_detect(
      raw_name,
      "ol_new_late_w_ol_new_late|ol_new_late_w_ol_old_late"
    )
  )

# Fail fast (defensive)
stopifnot(nrow(late_data_nature_HC) > 0)

# --------------------------------
# 2) Reshape to wide format
# --------------------------------

## Early
early_wide <- early_data_nature_HC %>%
  mutate(
    contrast = case_when(
      str_detect(raw_name, "ol_new_early_w_ol_new_early") ~ "new_new",
      str_detect(raw_name, "ol_new_early_w_ol_old_early") ~ "new_old"
    )
  ) %>%
 group_by(ID, Hall, contrast) %>%
summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop") %>%
pivot_wider(
  names_from  = contrast,
  values_from = RSA
)

# Optional sanity check
stopifnot(all(c("new_new", "new_old") %in% colnames(early_wide)))

## Late
late_wide <- late_data_nature_HC %>%
  mutate(
    contrast = case_when(
      str_detect(raw_name, "ol_new_late_w_ol_new_late") ~ "new_new",
      str_detect(raw_name, "ol_new_late_w_ol_old_late") ~ "new_old"
    )
  ) %>%
 
group_by(ID, Hall, contrast) %>%
summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop") %>%
pivot_wider(
  names_from  = contrast,
  values_from = RSA
)

# Optional sanity check
stopifnot(all(c("new_new", "new_old") %in% colnames(late_wide)))

```
paired t-tests (for both Early & Late), Cohen's d, CI's 95%...
```{r}

  full_paired_tests_HC <- function(data, LearningStage_label) {

  data %>%
    group_by(Hall) %>%
    summarise(
      n = sum(!is.na(new_new) & !is.na(new_old)),

      mean_new_new = mean(new_new, na.rm = TRUE),
      mean_new_old = mean(new_old, na.rm = TRUE),

      mean_diff = mean(new_new - new_old, na.rm = TRUE),
      sd_diff   = sd(new_new - new_old, na.rm = TRUE),

      t_test = list(t.test(new_new, new_old, paired = TRUE)),
      .groups = "drop"
    ) %>%

    mutate(
      t_value = sapply(t_test, function(x) unname(x$statistic)),
      df      = sapply(t_test, function(x) unname(x$parameter)),
      p_value = sapply(t_test, function(x) x$p.value),

      ci_low  = sapply(t_test, function(x) x$conf.int[1]),
      ci_high = sapply(t_test, function(x) x$conf.int[2]),

      # Cohen's dz for paired samples
      cohens_d = ifelse(sd_diff == 0, NA, mean_diff / sd_diff),

      # Standard error of the difference
      se_diff = sd_diff / sqrt(n),

      # APA/JASP-style p-values
      pValueFinal = case_when(
        p_value < .001 ~ "< .001",
        TRUE ~ paste0(
          "= ",
          sub("^0\\.", ".", sprintf("%.3f", round(p_value, 3)))
        )
      ),

      sig_value = case_when(
        p_value < .001 ~ "***",
        p_value < .01  ~ "**",
        p_value < .05  ~ "*",
        TRUE           ~ ""
      ),

      leStage = LearningStage_label
    ) %>%

    # Optional rounding for cleaner tables
    mutate(
      across(
        c(
          mean_new_new,
          mean_new_old,
          mean_diff,
          sd_diff,
          se_diff,
          t_value,
          ci_low,
          ci_high,
          cohens_d
        ),
        ~ round(.x, 3)
      )
    ) %>%

    select(
      leStage,
      Hall,
      n,
      mean_new_new,
      mean_new_old,
      mean_diff,
      sd_diff,
      se_diff,
      t_value,
      df,
      pValueFinal,
      ci_low,
      ci_high,
      cohens_d,
      sig_value
    )
}


```
Data table
```{r}
early_results_nature_HC<- full_paired_tests_HC(early_wide, "Early")
late_results_nature_HC  <- full_paired_tests_HC(late_wide,  "Late")

final_results_nature_HC <- bind_rows(early_results_nature_HC, late_results_nature_HC)
```
#  2b.2) RSC (EARLY+LATE)

```{r}

# 1) Filter data
## Early
early_data_nature_RSC <- long_parsed %>%
  filter(
    Overlappicity == "OL",
    learningStage == "Early",
    ROI == "RSC",
    str_detect(
      raw_name,
      "ol_new_early_w_ol_new_early|ol_new_early_w_ol_old_early"
    )
  )

# Fail fast (defensive)
stopifnot(nrow(early_data_nature_RSC) > 0)

## Late
late_data_nature_RSC <- long_parsed %>%
  filter(
    Overlappicity == "OL",
    learningStage == "Late",
    ROI == "RSC",
    str_detect(
      raw_name,
      "ol_new_late_w_ol_new_late|ol_new_late_w_ol_old_late"
    )
  )

# Fail fast (defensive)
stopifnot(nrow(late_data_nature_RSC) > 0)

# --------------------------------
# 2) Reshape to wide format
# --------------------------------

## Early
early_wide <- early_data_nature_RSC %>%
  mutate(
    contrast = case_when(
      str_detect(raw_name, "ol_new_early_w_ol_new_early") ~ "new_new",
      str_detect(raw_name, "ol_new_early_w_ol_old_early") ~ "new_old"
    )
  ) %>%
  
 group_by(ID, Hall, contrast) %>%
summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop") %>%
pivot_wider(
  names_from  = contrast,
  values_from = RSA
)

# Optional sanity check
stopifnot(all(c("new_new", "new_old") %in% colnames(early_wide)))

## Late
late_wide <- late_data_nature_RSC %>%
  mutate(
    contrast = case_when(
      str_detect(raw_name, "ol_new_late_w_ol_new_late") ~ "new_new",
      str_detect(raw_name, "ol_new_late_w_ol_old_late") ~ "new_old"
    )
  ) %>%
  
group_by(ID, Hall, contrast) %>%
summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop") %>%
pivot_wider(
  names_from  = contrast,
  values_from = RSA
)

# Optional sanity check
stopifnot(all(c("new_new", "new_old") %in% colnames(late_wide)))

```
paired t-tests (for both Early & Late), Cohen's d, CI's 95%...
```{r}

  full_paired_tests_RSC <- function(data, LearningStage_label) {

  data %>%
    group_by(Hall) %>%
    summarise(
      n = sum(!is.na(new_new) & !is.na(new_old)),

      mean_new_new = mean(new_new, na.rm = TRUE),
      mean_new_old = mean(new_old, na.rm = TRUE),

      mean_diff = mean(new_new - new_old, na.rm = TRUE),
      sd_diff   = sd(new_new - new_old, na.rm = TRUE),

      t_test = list(t.test(new_new, new_old, paired = TRUE)),
      .groups = "drop"
    ) %>%

    mutate(
      t_value = sapply(t_test, function(x) unname(x$statistic)),
      df      = sapply(t_test, function(x) unname(x$parameter)),
      p_value = sapply(t_test, function(x) x$p.value),

      ci_low  = sapply(t_test, function(x) x$conf.int[1]),
      ci_high = sapply(t_test, function(x) x$conf.int[2]),

      # Cohen's dz for paired samples
      cohens_d = ifelse(sd_diff == 0, NA, mean_diff / sd_diff),

      # Standard error of the difference
      se_diff = sd_diff / sqrt(n),

      # APA/JASP-style p-values
      pValueFinal = case_when(
        p_value < .001 ~ "< .001",
        TRUE ~ paste0(
          "= ",
          sub("^0\\.", ".", sprintf("%.3f", round(p_value, 3)))
        )
      ),

      sig_value = case_when(
        p_value < .001 ~ "***",
        p_value < .01  ~ "**",
        p_value < .05  ~ "*",
        TRUE           ~ ""
      ),

      leStage = LearningStage_label
    ) %>%

    # Optional rounding for cleaner tables
    mutate(
      across(
        c(
          mean_new_new,
          mean_new_old,
          mean_diff,
          sd_diff,
          se_diff,
          t_value,
          ci_low,
          ci_high,
          cohens_d
        ),
        ~ round(.x, 3)
      )
    ) %>%

    select(
      leStage,
      Hall,
      n,
      mean_new_new,
      mean_new_old,
      mean_diff,
      sd_diff,
      se_diff,
      t_value,
      df,
      pValueFinal,
      ci_low,
      ci_high,
      cohens_d,
      sig_value
    )
}


```
Data table
```{r}
early_results_nature_RSC <- full_paired_tests_RSC(early_wide, "Early")
late_results_nature_RSC  <- full_paired_tests_RSC(late_wide,  "Late")

final_results_nature_RSC <- bind_rows(early_results_nature_RSC, late_results_nature_RSC)
```
#  2b.3) RSComplex (EARLY+LATE)

```{r}

# 1) Filter data
## Early
early_data_nature_RSComplex <- long_parsed %>%
  filter(
    Overlappicity == "OL",
    learningStage == "Early",
    ROI == "Complex",  # keep INTERNAL name
    str_detect(
      raw_name,
      "ol_new_early_w_ol_new_early|ol_new_early_w_ol_old_early"
    )
  )

# ✅ Fail fast
stopifnot(nrow(early_data_nature_RSComplex) > 0)

## Late
late_data_nature_RSComplex <- long_parsed %>%
  filter(
    Overlappicity == "OL",
    learningStage == "Late",
    ROI == "Complex",  # keep INTERNAL name
    str_detect(
      raw_name,
      "ol_new_late_w_ol_new_late|ol_new_late_w_ol_old_late"
    )
  )

# ✅ Fail fast
stopifnot(nrow(late_data_nature_RSComplex) > 0)

# --------------------------------
# 2) Reshape to wide format
# --------------------------------

## Early
early_wide <- early_data_nature_RSComplex %>%
  mutate(
    contrast = case_when(
      str_detect(raw_name, "ol_new_early_w_ol_new_early") ~ "new_new",
      str_detect(raw_name, "ol_new_early_w_ol_old_early") ~ "new_old"
    )
  ) %>%
  
  group_by(ID, Hall, contrast) %>%
summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop") %>%
pivot_wider(
  names_from  = contrast,
  values_from = RSA
)

# ✅ Sanity check
stopifnot(all(c("new_new", "new_old") %in% colnames(early_wide)))

## Late
late_wide <- late_data_nature_RSComplex %>%
  mutate(
    contrast = case_when(
      str_detect(raw_name, "ol_new_late_w_ol_new_late") ~ "new_new",
      str_detect(raw_name, "ol_new_late_w_ol_old_late") ~ "new_old"
    )
  ) %>%
  
  group_by(ID, Hall, contrast) %>%
summarise(RSA = mean(RSA, na.rm = TRUE), .groups = "drop") %>%
pivot_wider(
  names_from  = contrast,
  values_from = RSA
)

# ✅ Sanity check
stopifnot(all(c("new_new", "new_old") %in% colnames(late_wide)))

```
paired t-tests (for both Early & Late), Cohen's d, CI's 95%...
```{r}

  full_paired_tests_RSComplex <- function(data, LearningStage_label) {

  data %>%
    group_by(Hall) %>%
    summarise(
      n = sum(!is.na(new_new) & !is.na(new_old)),

      mean_new_new = mean(new_new, na.rm = TRUE),
      mean_new_old = mean(new_old, na.rm = TRUE),

      mean_diff = mean(new_new - new_old, na.rm = TRUE),
      sd_diff   = sd(new_new - new_old, na.rm = TRUE),

      t_test = list(t.test(new_new, new_old, paired = TRUE)),
      .groups = "drop"
    ) %>%

    mutate(
      t_value = sapply(t_test, function(x) unname(x$statistic)),
      df      = sapply(t_test, function(x) unname(x$parameter)),
      p_value = sapply(t_test, function(x) x$p.value),

      ci_low  = sapply(t_test, function(x) x$conf.int[1]),
      ci_high = sapply(t_test, function(x) x$conf.int[2]),

      # Cohen's dz for paired samples
      cohens_d = ifelse(sd_diff == 0, NA, mean_diff / sd_diff),

      # Standard error of the difference
      se_diff = sd_diff / sqrt(n),

      # APA/JASP-style p-values
      pValueFinal = case_when(
        p_value < .001 ~ "< .001",
        TRUE ~ paste0(
          "= ",
          sub("^0\\.", ".", sprintf("%.3f", round(p_value, 3)))
        )
      ),

      sig_value = case_when(
        p_value < .001 ~ "***",
        p_value < .01  ~ "**",
        p_value < .05  ~ "*",
        TRUE           ~ ""
      ),

      leStage = LearningStage_label
    ) %>%

    # Optional rounding for cleaner tables
    mutate(
      across(
        c(
          mean_new_new,
          mean_new_old,
          mean_diff,
          sd_diff,
          se_diff,
          t_value,
          ci_low,
          ci_high,
          cohens_d
        ),
        ~ round(.x, 3)
      )
    ) %>%

    select(
      leStage,
      Hall,
      n,
      mean_new_new,
      mean_new_old,
      mean_diff,
      sd_diff,
      se_diff,
      t_value,
      df,
      pValueFinal,
      ci_low,
      ci_high,
      cohens_d,
      sig_value
    )
}

```
Data table

```{r}
early_results_nature_RSComplex <- full_paired_tests_RSComplex(early_wide, "Early")
late_results_nature_RSComplex  <- full_paired_tests_RSComplex(late_wide,  "Late")

final_results_nature_RSComplex <- bind_rows(early_results_nature_RSComplex, late_results_nature_RSComplex)
```
# Cleaning data: all 3 ROIs together 4 plotting (EARLY)

a) Data frame rearranged
```{r}
# add a ROI column to each "final_results"table
hipp_results    <- final_results_nature_HC    %>% mutate(ROI = "Hippocampus")
rsc_results     <- final_results_nature_RSC     %>% mutate(ROI = "RSC")
complex_results <- final_results_nature_RSComplex %>% mutate(ROI = "Complex")

# All ROI final results tables combined
final_results_allROIs <- bind_rows(
  hipp_results,
  rsc_results,
  complex_results
)
```
b) Prepare and define X-Axis order

```{r}
# 1) Filter Early and construct a custom x-axis factor

# library(dplyr)

early_plot <- final_results_allROIs %>%
  filter(leStage == "Early") %>%
  mutate(
    ROI = dplyr::recode(ROI,
                        "Hippocampus" = "HC",
                        "Complex"     = "RSComplex"),
    x_label = paste(Hall, ROI, sep = "_")
  )

# Factor levels
early_plot <- early_plot %>%
  mutate(
    x_label = factor(
      x_label,
      levels = c(
        "Cue period_HC", "Cue period_RSC", "Cue period_RSComplex",
        "1st hallway_HC", "1st hallway_RSC", "1st hallway_RSComplex",
        "2nd hallway_HC", "2nd hallway_RSC", "2nd hallway_RSComplex",
        "Goal_HC", "Goal_RSC", "Goal_RSComplex"
      )
    )
  )
```

****************************************************
# PLOTTING

## Early

Dot plots
****************************************************

```{r}

# # --------------------------------
# # Collapse across ROI within Hall (EARLY)
# # --------------------------------

early_plot_collapsed <- final_results_allROIs %>%
  filter(leStage == "Early") %>%
  mutate(
  ROI_label = dplyr::recode(ROI, "Complex" = "RSComplex")
)

y_breaks <- seq(-0.01, 0.06, by = 0.01)

roi_colors_early <- c(
  "Hippocampus" = "#E69F00", # muted orange
  "RSC"         = "#7B3294",  # deep purple
  "RSComplex"   = "#0072B2" # muted blue
)

dodge_width <- 0.6 # separation between ROIs/dots
star_dodge <- 0.60 # separation between stars

dotPlotEarly <- ggplot(
  early_plot_collapsed,
  aes(
    x     = Hall,
    y     = mean_diff,
    color = ROI_label,
    group = ROI_label
  )
) +

  # Error bars
  geom_errorbar(
    aes(
      ymin = mean_diff - se_diff,
      ymax = mean_diff + se_diff
    ),
    width = 0.15,
    linewidth = 0.6,
    color = "black",
    position = position_dodge(width = dodge_width)
  ) +

  # Points
  geom_point(
    size = 3.5,
    position = position_dodge(width = dodge_width)
  ) +

  # Significance stars 
  geom_text(
    aes(
      label = sig_value,
      y = mean_diff + se_diff + 0.003 # vertical position
    ),
    position = position_dodge(width = star_dodge), # horizontal position
    color = "black",
    size = 5, # star size
    fontface = "bold"
  ) +

  # Zero reference line
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.6,
    color = "grey40"
  ) +

  # Y axis
  # scale_y_continuous(
  #   limits = c(-0.01, 0.08),
  #   breaks = y_breaks,
  #   labels = scales::number_format(accuracy = 0.01),
  #   expand = c(0, 0)
  # ) +
  
  # Y axis: zooming in
scale_y_continuous(
  breaks = y_breaks,
  labels = scales::number_format(accuracy = 0.01),
  expand = c(0, 0)
) +

coord_cartesian(
  ylim = c(-0.01, 0.06)
) +

  # X axis with 4 ticks
  scale_x_discrete(
    limits = c("Cue period", "1st hallway", "2nd hallway", "Goal")
  ) +

  # Colors
  scale_color_manual(
    values = roi_colors_early,
    name   = "ROI"
  ) +

  labs(
    x = "Hall",
    y = "(OL New+New) − (OL New+Old)"
  ) +

  # Panel lettering "A"
  annotate(
  "text",
  x = 0.5,
  y = Inf,
  vjust = 1.2,
  label = "A",
  fontface = "bold",
  size = 6,
  hjust = 0
) +

  theme_classic(base_size = 14) +
  theme(
    axis.text.x  = element_text(color = "black"),
    axis.text.y  = element_text(color = "black"),
    axis.title.y = element_text(face = "bold", size = 12),
    axis.line    = element_line(linewidth = 0.8),
    axis.ticks   = element_line(linewidth = 0.8),
    #legend.position = "none"
  )
  

# Debug required in R Markdown
print(dotPlotEarly)

# Save graph
ggsave(
  "Early_Repulsion_t-tests_collapsed_halls_Final_A_Y-axis-Maxx.tiff",
  plot = dotPlotEarly,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)



```

****************************************************
Dot plot (LATE)  collapsed halls- v2 (02/05/2026) **needs to fix the asterisks
****************************************************
```{r}
# --------------------------------
# Collapse across ROI within Hall (LATE)
# --------------------------------

late_plot_collapsed <- final_results_allROIs %>%
  filter(leStage == "Late") %>%
  mutate(
    ROI_label = dplyr::recode(ROI, "Complex" = "RSComplex")
  )

y_breaks <- seq(-0.01, 0.06, by = 0.01)

roi_colors_late <- c(
  "Hippocampus" = "#FF7F00",  # bright orange
  "RSC"         = "#8B00FF",  # ultraviolet
  "RSComplex"   = "#007FFF"   # bright blue
)

dodge_width <- 0.6
star_dodge <- 0.6 # separation between stars

dotPlotLate <- ggplot(
  late_plot_collapsed,
  aes(
    x     = Hall,
    y     = mean_diff,
    color = ROI_label,
    group = ROI_label
  )
) +

  # Error bars
  geom_errorbar(
    aes(
      ymin = mean_diff - se_diff,
      ymax = mean_diff + se_diff
    ),
    width = 0.15,
    linewidth = 0.6,
    color = "black",
    position = position_dodge(width = dodge_width)
  ) +

  # Points
  geom_point(
    size = 3.5,
    position = position_dodge(width = dodge_width)
  ) +

  # Significance stars (same logic as EARLY)
  geom_text(
    aes(
      label = sig_value,
      y     = mean_diff + se_diff + 0.003
    ),
    position = position_dodge(width = dodge_width),
    color = "black",
    size = 5,
    fontface = "bold"
  ) +

  # Zero reference line
  geom_hline(
    yintercept = 0,
    linetype   = "dashed",
    linewidth  = 0.6,
    color      = "grey40"
  ) +

  # Y axis
  # scale_y_continuous(
  #   limits = c(-0.01, 0.08),
  #   breaks = y_breaks,
  #   labels = scales::number_format(accuracy = 0.01),
  #   expand = c(0, 0)
  # ) +
  
  # Y axis
scale_y_continuous(
  breaks = y_breaks,
  labels = scales::number_format(accuracy = 0.01),
  expand = c(0, 0)
) +

coord_cartesian(
  ylim = c(-0.01, 0.06)
) +

  # ✅ X axis: 4 ticks only (same as EARLY)
  scale_x_discrete(
    limits = c("Cue period", "1st hallway", "2nd hallway", "Goal")
  ) +

  # Colors
  scale_color_manual(
    values = roi_colors_late,
    name   = "ROI"
  ) +

  labs(
    x = "Hall",
    y = "(OL New+New) − (OL New+Old)"
  ) +

 # Panel lettering "B"
  annotate(
  "text",
  x = 0.5,
  y = Inf,
  vjust = 1.2,
  label = "B",
  fontface = "bold",
  size = 6,
  hjust = 0
) +

  theme_classic(base_size = 14) +
  theme(
    axis.text.x  = element_text(color = "black"),
    axis.text.y  = element_text(color = "black"),
    axis.title.y = element_text(face = "bold", size = 12),
    axis.line    = element_line(linewidth = 0.8),
    axis.ticks   = element_line(linewidth = 0.8),
    #legend.position = "none"
  )

# Debug required in R Markdown
print(dotPlotLate)

# Save graph
ggsave(
  "Late_Repulsion_ttests_collapsed_Hall_Final_B_Y-axis-Maxx.tiff",
  plot = dotPlotLate,
  width = 7,
  height = 6,
  dpi = 600,
  compression = "lzw"
)

```

***************
# 3) CORRELATIONS (with no corrections)
***************
# a) OL Early

 HIPPOCAMPUS + RSC

```{r}

# library(dplyr)
# library(tidyr)
# library(ggplot2)
# library(stringr)

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("RSC", "Hippocampus"),
    str_detect(raw_name, "ol_new_early_w_ol_new_early_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches exactly)
# ------------------------------------------------------------
axis_order <- c(
  "Hippocampus Cue period",
  "Hippocampus 1st hallway",
  "Hippocampus 2nd hallway",
  "Hippocampus Goal",
  "RSC Cue period",
  "RSC 1st hallway",
  "RSC 2nd hallway",
  "RSC Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}

# Choose which p-values (RAW = p-mat or CORRECTED = p_adj_mat) to use for stars:
p_use <- p_mat  # or use p_adj_mat if enabled Benjamini–Hochberg (BH) correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df

```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )

```
## PLOTTING
# CORRELATION MATRIX HC + RSC OL EARLY
```{r}

# --- Palette prep (tone down viridis) ---
library(viridisLite)
library(colorspace)  # <-- use this for desaturate()
library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_OL_Early_HC_RSC <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  # scale_x_discrete(limits = present) +
  # scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).
  

  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +

scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white", # keeps the masked triangle white
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
  labs(
  x = NULL,
  y = NULL,
  title = "HC + RSC OL EARLY"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_OL_Early_HC_RSC)

# Save
ggsave("matrix_OL_Early_HC_RSC-upper.tiff",
       plot = matrix_OL_Early_HC_RSC,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```

### HIPPOCAMPUS + RSC

# OL Late

 HC + RSC OL LATE MATRIX
```{r}

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("RSC", "Hippocampus"),
    str_detect(raw_name, "ol_new_late_w_ol_new_late_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches exactly)
# ------------------------------------------------------------
axis_order <- c(
  "Hippocampus Cue period",
  "Hippocampus 1st hallway",
  "Hippocampus 2nd hallway",
  "Hippocampus Goal",
  "RSC Cue period",
  "RSC 1st hallway",
  "RSC 2nd hallway",
  "RSC Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df

```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX HC + RSC OL LATE
```{r}

# --- Palette prep (tone down viridis) ---
# library(viridisLite)
# library(colorspace)  # <-- use this for desaturate()
# library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_OL_Late_HC_RSC <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  # scale_x_discrete(limits = present) +
  # scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).
  
  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +

scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
   na.value = "white", # keeps the masked triangle white
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
  labs(
  x = NULL,
  y = NULL,
  title = "HC + RSC OL LATE"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_OL_Late_HC_RSC)

# EXPORT AT JOURNAL RESOLUTION
ggsave("matrix_OL_Late_HC_RSC-upper.tiff",
       plot = matrix_OL_Late_HC_RSC,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```
### HIPPOCAMPUS + RSCOMPLEX

# OL Early

 HC + RSComplex OL EARLY MATRIX

```{r}

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "Hippocampus"),
    str_detect(raw_name, "ol_new_early_w_ol_new_early_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches exactly the heatmap_wide df labels)
# ------------------------------------------------------------
axis_order <- c(
  "Hippocampus Cue period",
  "Hippocampus 1st hallway",
  "Hippocampus 2nd hallway",
  "Hippocampus Goal",
  "Complex Cue period",
  "Complex 1st hallway",
  "Complex 2nd hallway",
  "Complex Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df

```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX HC + RSComplex OL EARLY
```{r}

# --- Palette prep (tone down viridis) ---
# library(viridisLite)
# library(colorspace)  # <-- use this for desaturate()
# library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_OL_Early_HC_RSComplex <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  scale_x_discrete(limits = present) +
  scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).
  
  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white", # keeps the masked triangle white,
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
#   scale_x_discrete(
#   limits = present,
#   labels = function(x) sub("^Complex", "RSComplex", x)  # rename df labels only for display
# ) +
# scale_y_discrete(
#   limits = present,
#   labels = function(y) sub("^Complex", "RSComplex", y)  # rename df labels 
# ) +
  
  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +

scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
  labs(
  x = NULL,
  y = NULL,
  title = "HC + RSComplex OL EARLY"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_OL_Early_HC_RSComplex)


# EXPORT AT JOURNAL RESOLUTION
ggsave("matrix_OL_Early_HC_RSComplex-upper.tiff",
       plot = matrix_OL_Early_HC_RSComplex,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```
### HIPPOCAMPUS + RSComplex

# OL Late

 HC + RSComplex OL LATE MATRIX
```{r}

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "Hippocampus"),
    str_detect(raw_name, "ol_new_late_w_ol_new_late_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches heatmap_wide df exactly)
# ------------------------------------------------------------
axis_order <- c(
  "Hippocampus Cue period",
  "Hippocampus 1st hallway",
  "Hippocampus 2nd hallway",
  "Hippocampus Goal",
  "Complex Cue period",
  "Complex 1st hallway",
  "Complex 2nd hallway",
  "Complex Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df

```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX HC + RSComplex OL LATE
```{r}

# --- Palette prep (tone down viridis) ---
# library(viridisLite)
# library(colorspace)  # <-- use this for desaturate()
# library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_OL_Late_HC_RSComplex <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  scale_x_discrete(limits = present) +
  scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).
  
  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white",
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
  
#   scale_x_discrete(
#   limits = present,
#   labels = function(x) sub("^Complex", "RSComplex", x)  # rename df labels only for display
# ) +
# scale_y_discrete(
#   limits = present,
#   labels = function(y) sub("^Complex", "RSComplex", y)  # rename df labels 
# ) +
  
  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +

scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
  labs(
  x = NULL,
  y = NULL,
  title = "HC + RSComplex OL LATE"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_OL_Late_HC_RSComplex)

# EXPORT AT JOURNAL RESOLUTION
ggsave("HC + RSComplex OL late_matrix-upper.tiff",
       plot = matrix_OL_Late_HC_RSComplex,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```
##  RSC + RSComplex

# OL Early
  
 # RSC + RSComplex OL EARLY
 
```{r}

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "RSC"),
    str_detect(raw_name, "ol_new_early_w_ol_new_early_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches exactly the heatmap_wide df labels)
# ------------------------------------------------------------
axis_order <- c(
  "RSC Cue period",
  "RSC 1st hallway",
  "RSC 2nd hallway",
  "RSC Goal",
  "Complex Cue period",
  "Complex 1st hallway",
  "Complex 2nd hallway",
  "Complex Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df

```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX RSC + RSComplex OL EARLY
```{r}

# --- Palette prep (tone down viridis) ---
# library(viridisLite)
# library(colorspace)  # <-- use this for desaturate()
# library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_OL_Early_RSC_RSComplex <-ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  scale_x_discrete(limits = present) +
  scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).

  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white",
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
#   scale_x_discrete(
#   limits = present,
#   labels = function(x) sub("^Complex", "RSComplex", x)  # rename df labels only for display
# ) +
# scale_y_discrete(
#   limits = present,
#   labels = function(y) sub("^Complex", "RSComplex", y)  # rename df labels 
# ) +
  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +

scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
  labs(
  x = NULL,
  y = NULL,
  title = "RSC + RSComplex OL EARLY"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_OL_Early_RSC_RSComplex)
# EXPORT AT JOURNAL RESOLUTION
ggsave("matrix_OL_Early_RSC_RSComplex-upper.tiff",
       plot = matrix_OL_Early_RSC_RSComplex,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```

### RSC + RSComplex
# OL  Late

 RSC + RSComplex OL LATE MATRIX
```{r}

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "RSC"),
    str_detect(raw_name, "ol_new_late_w_ol_new_late_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches heatmap_wide df exactly)
# ------------------------------------------------------------
axis_order <- c(
  "RSC Cue period",
  "RSC 1st hallway",
  "RSC 2nd hallway",
  "RSC Goal",
  "Complex Cue period",
  "Complex 1st hallway",
  "Complex 2nd hallway",
  "Complex Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df

```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX RSC + RSComplex OL LATE
```{r}

# --- Palette prep (tone down viridis) ---
#  library(viridisLite)
# library(colorspace)  # <-- use this for desaturate()
# library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_OL_Late_RSC_RSComplex <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  scale_x_discrete(limits = present) +
  scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).
  
  # scale_fill_gradient2( # Sets the color scale for fill
  #   low = "white", # for a 0-centered diverging scale with a few negative values, set a light color for negatives
  #   high = "blue",
  #   na.value = "white",
  #   midpoint = 0, # centers the color scale at 0 (useful for correlations)
  #   limits = c(-0.2, 1),
  #   name = expression(italic(r)) # legend title
  # )
  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white",
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
  
#   scale_x_discrete(
#   limits = present,
#   labels = function(x) sub("^Complex", "RSComplex", x)  # rename df labels only for display
# ) +
# scale_y_discrete(
#   limits = present,
#   labels = function(y) sub("^Complex", "RSComplex", y)  # rename df labels 
# ) +
  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +

scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  labs(
  x = NULL,
  y = NULL,
  title = "RSC + RSComplex OL LATE"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_OL_Late_RSC_RSComplex)

# EXPORT AT JOURNAL RESOLUTION
ggsave("matrix_OL_Late_RSC_RSComplex-upper.tiff",
       plot = matrix_OL_Late_RSC_RSComplex,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```

************************
# 3. b) NON-OVERLAPPING 

## NOL EARLY


 HC + RSC 

```{r}
# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("RSC", "Hippocampus"),
    str_detect(raw_name, "nol_new_early_w_nol_new_early_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = paste(ROI, Hall) 
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches exactly)
# ------------------------------------------------------------
axis_order <- c(
  "Hippocampus Cue period",
  "Hippocampus 1st hallway",
  "Hippocampus 2nd hallway",
  "Hippocampus Goal",
  "RSC Cue period",
  "RSC 1st hallway",
  "RSC 2nd hallway",
  "RSC Goal"
)

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df

```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX HC + RSC NOL EARLY
```{r}

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_NOL_Early_HC_RSC <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  # scale_x_discrete(limits = present) +
  # scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).
  
scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("^Hippocampus", "HC", x)
    x <- gsub("^RSC", "RSC", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
  
  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white",
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
  labs(
  x = NULL,
  y = NULL,
  title = "HC + RSC NOL EARLY"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_NOL_Early_HC_RSC)

# EXPORT AT JOURNAL RESOLUTION
ggsave("matrix_NOL_Early_HC_RSC-upper.tiff",
       plot = matrix_NOL_Early_HC_RSC,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```

### HIPPOCAMPUS + RSC

 HC + RSC NOL LATE MATRIX
```{r}

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("RSC", "Hippocampus"),
    str_detect(raw_name, "nol_new_late_w_nol_new_late_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches exactly)
# ------------------------------------------------------------
axis_order <- c(
  "Hippocampus Cue period",
  "Hippocampus 1st hallway",
  "Hippocampus 2nd hallway",
  "Hippocampus Goal",
  "RSC Cue period",
  "RSC 1st hallway",
  "RSC 2nd hallway",
  "RSC Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df


```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX HC + RSC NOL LATE
```{r}

# --- Palette prep (tone down viridis) ---
#  library(viridisLite)
# library(colorspace)  # <-- use this for desaturate()
# library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_NOL_Late_HC_RSC <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  # scale_x_discrete(limits = present) +
  # scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).
  # 
  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("^Hippocampus", "HC", x)
    x <- gsub("^RSC", "RSC", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
  scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("^Hippocampus", "HC", x)
    x <- gsub("^RSC", "RSC", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white",
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
  labs(
  x = NULL,
  y = NULL,
  title = "HC + RSC NOL LATE"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_NOL_Late_HC_RSC)

# EXPORT AT JOURNAL RESOLUTION
ggsave("matrix_NOL_Late_HC_RSC-upper.tiff",
       plot = matrix_NOL_Late_HC_RSC,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```
### HIPPOCAMPUS + RSCOMPLEX

 HC + RSComplex NOL EARLY MATRIX

```{r}

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "Hippocampus"),
    str_detect(raw_name, "nol_new_early_w_nol_new_early_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches exactly the heatmap_wide df labels)
# ------------------------------------------------------------
axis_order <- c(
  "Hippocampus Cue period",
  "Hippocampus 1st hallway",
  "Hippocampus 2nd hallway",
  "Hippocampus Goal",
  "Complex Cue period",
  "Complex 1st hallway",
  "Complex 2nd hallway",
  "Complex Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df

```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX HC + RSComplex NOL EARLY
```{r}

# --- Palette prep (tone down viridis) ---
#  library(viridisLite)
# library(colorspace)  # <-- use this for desaturate()
# library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_NOL_Early_HC_RSComplex <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  scale_x_discrete(limits = present) +
  scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).
  
  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white",
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
#   scale_x_discrete(
#   limits = present,
#   labels = function(x) sub("^Complex", "RSComplex", x)  # rename df labels only for display
# ) +
# scale_y_discrete(
#   limits = present,
#   labels = function(y) sub("^Complex", "RSComplex", y)  # rename df labels 
# ) +
  
  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +

scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
  labs(
  x = NULL,
  y = NULL,
  title = "HC + RSComplex NOL EARLY"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_NOL_Early_HC_RSComplex)

# EXPORT AT JOURNAL RESOLUTION
ggsave("matrix_NOL_Early_HC_RSComplex-upper.tiff",
       plot = matrix_NOL_Early_HC_RSComplex,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```

### HIPPOCAMPUS + RSComplex

 HC + RSComplex NOL LATE MATRIX
```{r}

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "Hippocampus"),
    str_detect(raw_name, "nol_new_late_w_nol_new_late_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches heatmap_wide df exactly)
# ------------------------------------------------------------
axis_order <- c(
  "Hippocampus Cue period",
  "Hippocampus 1st hallway",
  "Hippocampus 2nd hallway",
  "Hippocampus Goal",
  "Complex Cue period",
  "Complex 1st hallway",
  "Complex 2nd hallway",
  "Complex Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df


```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX HC + RSComplex NOL LATE
```{r}

# --- Palette prep (tone down viridis) ---
#  library(viridisLite)
# library(colorspace)  # <-- use this for desaturate()
# library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_NOL_Late_HC_RSComplex <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  scale_x_discrete(limits = present) +
  scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).
  
  # scale_fill_gradient2( # Sets the color scale for fill
  #   low = "white", # for a 0-centered diverging scale with a few negative values, set a light color for negatives
  #   high = "blue",
  #   na.value = "white",
  #   midpoint = 0, # centers the color scale at 0 (useful for correlations)
  #   limits = c(-0.2, 1),
  #   name = expression(italic(r)) # legend title
  # )
  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white",
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
  
#   scale_x_discrete(
#   limits = present,
#   labels = function(x) sub("^Complex", "RSComplex", x)  # rename df labels only for display
# ) +
# scale_y_discrete(
#   limits = present,
#   labels = function(y) sub("^Complex", "RSComplex", y)  # rename df labels 
# ) +
  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +

scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Hippocampus", "HC", x)
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
  labs(
  x = NULL,
  y = NULL,
  title = "HC + RSComplex NOL LATE"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_NOL_Late_HC_RSComplex)

# EXPORT AT JOURNAL RESOLUTION
ggsave("HC + RSComplex NOL late_matrix-upper.tiff",
       plot = matrix_NOL_Late_HC_RSComplex,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```
##  RSC + RSComplex
  
 # RSC + RSComplex NOL EARLY
 
```{r}

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "RSC"),
    str_detect(raw_name, "nol_new_early_w_nol_new_early_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches exactly the heatmap_wide df labels)
# ------------------------------------------------------------
axis_order <- c(
  "RSC Cue period",
  "RSC 1st hallway",
  "RSC 2nd hallway",
  "RSC Goal",
  "Complex Cue period",
  "Complex 1st hallway",
  "Complex 2nd hallway",
  "Complex Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df

```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX RSC + RSComplex NOL EARLY
```{r}

# --- Palette prep (tone down viridis) ---
# library(viridisLite)
# library(colorspace)  # <-- use this for desaturate()
# library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_NOL_Early_RSC_RSComplex <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  scale_x_discrete(limits = present) +
  scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).
  
  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white",
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
#   scale_x_discrete(
#   limits = present,
#   labels = function(x) sub("^Complex", "RSComplex", x)  # rename df labels only for display
# ) +
# scale_y_discrete(
#   limits = present,
#   labels = function(y) sub("^Complex", "RSComplex", y)  # rename df labels 
# ) +
  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +

scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
  labs(
  x = NULL,
  y = NULL,
  title = "RSC + RSComplex NOL EARLY"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_NOL_Early_RSC_RSComplex)

# EXPORT AT JOURNAL RESOLUTION
ggsave("RSC + RSComplex NOL Early_matrix-upper.tiff",
       plot = matrix_NOL_Early_RSC_RSComplex,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```

### RSC + RSComplex

 RSC + RSComplex NOL LATE MATRIX
```{r}

# --------------------------------
# 1) Filter + clean strings
# --------------------------------
heatmapFilter <- long_parsed %>%
  filter(
    ROI %in% c("Complex", "RSC"),
    str_detect(raw_name, "nol_new_late_w_nol_new_late_")
  ) %>%
  group_by(ID, ROI, Hall) %>%
  mutate(
    # normalize the Hall names to avoid hidden mismatches
    Hall = str_squish(Hall),
    Hall = factor(Hall, levels = c("Cue period", "1st hallway", "2nd hallway", "Goal")),
    ROI  = str_squish(ROI)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2) Create combined condition label (normalized)
# ------------------------------------------------------------
heatmapFilter <- heatmapFilter %>%
  mutate(
    Condition = str_squish(paste(ROI, Hall))
  )

# ------------------------------------------------------------
# 3) Pivot to wide format
# ------------------------------------------------------------
heatmap_wide <- heatmapFilter %>%
  select(ID, Condition, RSA) %>%
  pivot_wider(
    names_from  = Condition,
    values_from = RSA
  )

# ------------------------------------------------------------
# 4) Desired axis order (make sure it matches heatmap_wide df exactly)
# ------------------------------------------------------------
axis_order <- c(
  "RSC Cue period",
  "RSC 1st hallway",
  "RSC 2nd hallway",
  "RSC Goal",
  "Complex Cue period",
  "Complex 1st hallway",
  "Complex 2nd hallway",
  "Complex Goal"
)

# Sanity check
# print(setdiff(axis_order, colnames(heatmap_wide)))
# print(setdiff(colnames(heatmap_wide), axis_order))

# ------------------------------------------------------------
# 5) Compute correlation matrix (restricted + reordered)
# ------------------------------------------------------------
# Keep only the columns in axis_order (in that order), dropping any others
present <- intersect(axis_order, colnames(heatmap_wide))
if (length(present) != length(axis_order)) {
  warning("Some expected conditions are missing: ",
          paste(setdiff(axis_order, present), collapse = ", "))
}
corr_matrix <- heatmap_wide %>%
  select(all_of(present)) %>%
  cor(use = "pairwise.complete.obs")

# Reorder rows/cols to the desired axis_order (for present labels)
corr_matrix <- corr_matrix[present, present, drop = FALSE]

# ------------------------------------------------------------
# 5) Compute pairwise counts (n) and p-values for the matrix
# ------------------------------------------------------------
data_wide <- heatmap_wide %>% select(all_of(present))
X <- as.matrix(data_wide)

# Pairwise non-NA counts: n_ij = number of rows where both cols i and j are not NA
ind   <- !is.na(X)
n_mat <- t(ind) %*% ind    # matrix of pairwise sample sizes

# We already have r_mat = corr_matrix (reordered to [present, present])
r_mat <- corr_matrix

# Compute p-values from r and n (two-sided); handle r = ±1, n < 3 safely
p_mat <- matrix(NA_real_, nrow = nrow(r_mat), ncol = ncol(r_mat),
                dimnames = dimnames(r_mat))

for (i in seq_len(nrow(r_mat))) {
  for (j in seq_len(ncol(r_mat))) {
    r <- r_mat[i, j]
    n <- n_mat[i, j]
    if (!is.na(r) && !is.na(n) && n >= 3) {
      if (abs(r) < 1) {
        tval <- r * sqrt((n - 2) / (1 - r^2))
        pval <- 2 * pt(-abs(tval), df = n - 2)
      } else {
        # Perfect correlation with >= 3 points -> p ~ 0
        pval <- 0
      }
      p_mat[i, j] <- pval
    } else {
      p_mat[i, j] <- NA_real_
    }
  }
}


# Choose which p-values to use for stars:
p_use <- p_mat  # or use p_adj_mat if you enabled BH correction above

# Build significance stars
stars_mat <- matrix("", nrow = nrow(p_use), ncol = ncol(p_use),
                    dimnames = dimnames(p_use))
stars_mat[!is.na(p_use) & p_use < 0.001] <- "***"
stars_mat[!is.na(p_use) & p_use >= 0.001 & p_use < 0.01] <- "**"
stars_mat[!is.na(p_use) & p_use >= 0.01 & p_use < 0.05] <- "*"

# ------------------------------------------------------------
# 6)Convert correlation matrix (r mat) to long format and add p-values + stars
# ------------------------------------------------------------
heatmap_df <- as.data.frame(as.table(r_mat)) %>%
  rename(
    X_label = Var1,
    Y_label = Var2,
    cor     = Freq
  ) %>%
  mutate(
    # attach p and stars from matrices (matching by dimnames)
    p_value = as.vector(p_mat[cbind(match(X_label, rownames(p_mat)),
                                    match(Y_label, colnames(p_mat)))]),
    stars   = stars_mat[cbind(match(X_label, rownames(stars_mat)),
                              match(Y_label, colnames(stars_mat)))],
    # format r with 3 decimals and append stars
    label_r = sprintf("%.3f", cor),
    label   = ifelse(is.na(cor), "", paste0(label_r, stars))
  ) %>%
  mutate(
    X_label = factor(X_label, levels = present),
    Y_label = factor(Y_label, levels = present)
  )

# Freeze master copy
heatmap_master <- heatmap_df

```
## MASKING TRIANGLES
```{r}
## MASKING (UPPER OR LOWER) TRIANGLES

# Choose masking type
mask_type <- "lower"   # "none", "upper", "lower", "upper_diag", "lower_diag"

# Fresh plotting copy
heatmap_plot <- heatmap_master %>%
  mutate(
    x_num = as.numeric(X_label),
    y_num = as.numeric(Y_label)
  )

if(mask_type == "upper"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num < y_num, NA, cor),
      label = ifelse(x_num < y_num, "", label)
    )
}

if(mask_type == "lower"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num > y_num, NA, cor),
      label = ifelse(x_num > y_num, "", label)
    )
}

if(mask_type == "upper_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num <= y_num, NA, cor),
      label = ifelse(x_num <= y_num, "", label)
    )
}

if(mask_type == "lower_diag"){
  heatmap_plot <- heatmap_plot %>%
    mutate(
      cor   = ifelse(x_num >= y_num, NA, cor),
      label = ifelse(x_num >= y_num, "", label)
    )
}
 
## Remove values on the diagonal ONLY (self-correlations)
heatmap_plot <- heatmap_plot %>%
  mutate(
    label = ifelse(X_label == Y_label, NA_character_, label)  # NA suppresses geom_text
  )
# heatmap_plot <- heatmap_plot %>% * this one removes tile colors (blank)
#   mutate(
#     cor   = ifelse(X_label == Y_label, NA, cor),
#     label = ifelse(X_label == Y_label, "", label)
#   )
```
## PLOTTING
# CORRELATION MATRIX RSC + RSComplex NOL LATE
```{r}

# --- Palette prep (tone down viridis) ---
# library(viridisLite)
# library(colorspace)  # <-- use this for desaturate()
# library(scales)      # for alpha() and squish

cols      <- viridisLite::viridis(256, begin = 0.2, end = 0.95)
soft_cols <- scales::alpha(colorspace::desaturate(cols, amount = 0.1), 0.85) # amount in desaturate() → higher = more muted (e.g., 0.5, 0.7) / alpha alpha(..., 0.85) → lower = lighter (e.g., 0.8).
# amount: 0 = no change; 1 = fully gray

# Plot
matrix_NOL_Late_RSC_RSComplex <- ggplot(heatmap_plot, aes(x = X_label, y = Y_label, fill = cor)) +
  geom_tile(linewidth = 0) + # Draws each heatmap tile (a square) with no border if linewidth = 0.
  geom_text(aes(label = label), # Adds text labels on tiles using the label column (e.g., 0.347*).
            size = 3.3, # cell text size (e.g., 0.347*)
            fontface = "bold", # use "plain" to "unbold"
            color = "black") +
  scale_x_discrete(limits = present) +
  scale_y_discrete(limits = present) +  # Forces the axis order to match the vector (the list of condition names that actually exist in the data).

  
scale_fill_gradientn(
  colours = soft_cols, # reversed viridis color palette toned down (see above)
  limits  = c(0, 1), # or c(-0.2, 1) to show tiny negatives faintly
  oob     = scales::squish, # clip out-of-bounds to nearest limit
  na.value = "white",
  name    = expression(italic(r))
) + # virids is a continuous, color-blind friendly, color scale
  guides(
  fill = guide_colorbar(
    frame.colour   = "black",    # outline color of the colorbar
    frame.linewidth = 0.5,       # outline thickness
    ticks.colour   = "black"     # optional: tick marks on the bar
  )
) +
  
#   scale_x_discrete(
#   limits = present,
#   labels = function(x) sub("^Complex", "RSComplex", x)  # rename df labels only for display
# ) +
# scale_y_discrete(
#   limits = present,
#   labels = function(y) sub("^Complex", "RSComplex", y)  # rename df labels 
# ) +
  
  scale_x_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
scale_y_discrete(
  limits = present,
  labels = function(x) {
    x <- gsub("Complex", "RSComplex", x)
    x <- gsub("Cue period", "Cue", x)
    x <- gsub("1st hallway", "Hall1", x)
    x <- gsub("2nd hallway", "Hall2", x)
    x
  }
) +
  
  labs(
  x = NULL,
  y = NULL,
  title = "RSC + RSComplex NOL LATE"
) +
coord_fixed() +
theme_minimal(base_size = 13) +
theme(
  plot.title = element_text(
    hjust = 0.5,     # center title (0 = left, 1 = right)
    face = "bold",   # make title bold
    size = 12        # optional: adjust title size
  ),
  
  axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    face = "bold",
    color = "black"
  ),
  axis.text.y = element_text(
    face = "bold",
    color = "black"
  ),
  panel.grid = element_blank(),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(color = "black"),
  legend.key.height = unit(0.8, "cm"),
  legend.key.width  = unit(0.4, "cm")
)

print(matrix_NOL_Late_RSC_RSComplex)
# EXPORT AT JOURNAL RESOLUTION
ggsave("RSC + RSComplex NOL late_matrix-upper.tiff",
       plot = matrix_NOL_Late_RSC_RSComplex,
       width = 7,
       height = 6,
       dpi = 600,
       compression = "lzw")
```